<!-- ## Continuous Distributions -->


### One Parameter Distributions

<div class="tab">
  <button class="tablinks" onclick="openTab(event, '1pA')">Exponential</button>
  <button class="tablinks" onclick="openTab(event, '1pB')">Inv Exponential</button>
  <button class="tablinks" onclick="openTab(event, 'odC')">Single Parameter Pareto</button>
</div>


<div id="1pA" class="tabcontent">
  <span onclick="this.parentElement.style.display='none'" class="topright"><button>Hide</button></span>

<h3>Exponential</h3>

**Functions**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Name} & \text{Function} \\
\hline
  \small{\text{Parameter assumptions}} & \theta>0 \\
\hline
  \small{\text{Probability density}} & \frac{1}{\theta}e^{-x/\theta} \\
 ~~ \small{\text{function }} f(x)& \\
\hline
  \small{\text{Distribution function}} & 1-e^{-x/\theta} \\
   ~~F(x) & \\
\hline
 \textit{k}^{th}~\small{\text{raw moment}}  & \theta^k\Gamma(k+1) \\
  ~~\mathrm{E}[X^k]  & k>-1 \\
\hline
  VaR_p(x) & -\theta \ln (1-p) \\
\hline
  \small{\text{Limited Expected Value}} & \theta(1-e^{-x/\theta}) \\
  ~~\mathrm{E}[X\wedge x] & \\
\hline
\end{array}
\end{matrix}
$$

**`R` Commands**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Function Name} & \text{R Command} \\
\hline
  \small{\text{Density function}} & \text{dexp}(x=, rate=1/\theta) \\
\hline
  \small{\text{Distribution function}} & \text{pexp}(p=, rate=1/\theta) \\
\hline
  \small{\text{Quantile function}} & \text{qexp}(q=, rate=1/\theta) \\ 
\hline
  \small{\text{Random sampling function}} & \text{rexp}(n=, rate=1/\theta) \\
\hline
\end{array}
\end{matrix}
$$

**Illustrative Graph**

```{r echo=SHOW_PDF}
theta <- 100
X <- seq(from = 0, to = 1000, by = 1)
plot(x=X,y=dexp(X,rate=1/theta),type="l",
     ylab="Probability density",col="red",main="Exponential Distribution")
```

</div>

<div id="1pB" class="tabcontent">
  <span onclick="this.parentElement.style.display='none'" class="topright"><button>Hide</button></span>

<h3>Inverse Exponential</h3>


**Functions**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Name} & \text{Function} \\
    \hline
  \small{\text{Parameter assumptions}} & \theta>0 \\
\hline
  \small{\text{Probability density}} & \frac{\theta e^{-\theta/x}}{x^2} \\
 ~~ \small{\text{function }} f(x)& \\
\hline
  \small{\text{Distribution function}} & e^{-\theta/x} \\
   ~~F(x) & \\
\hline
  \textit{k}^{th}~\small{\text{raw moment}} & \theta^k\Gamma(1-k) \\
  ~~\mathrm{E}[X^k]  & k<1 \\
\hline
  \mathrm{E}[(X\wedge x)^k] & \theta^kG(1-k;\theta/x)+x^k (1 - e^{-\theta/x}) \\
\hline
\end{array}
\end{matrix}
$$


**`R` Commands**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Function Name} & \text{R Command} \\
\hline
  \small{\text{Density function}} & \text{dinvexp}(x=, scale=\theta) \\
\hline
  \small{\text{Distribution function}} & \text{pinvexp}(p=, scale=\theta) \\
\hline
  \small{\text{Quantile function}} & \text{qinvexp}(q=, scale=\theta) \\ 
\hline
  \small{\text{Random sampling function}} & \text{rinvexp}(n=, scale=\theta) \\
\hline
\end{array}
\end{matrix}
$$

**Illustrative Graph**

```{r echo=SHOW_PDF}
theta <- 0.01
X <- seq(from = 0, to = 1000, by = 1)
plot(x=X,y=dinvexp(X, rate =  1/theta),type="l",
     ylab="Probability density",col="red",main="Inverse Exponential Distribution")
```
</div>


<div id="odC" class="tabcontent">
  <span onclick="this.parentElement.style.display='none'" class="topright"><button>Hide</button></span>

<h3>Single Parameter Pareto</h3>

**Functions**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Name} & \text{Function} \\
\hline
  \small{\text{Parameter assumptions}} & \theta~\text{is known},~x>\theta, \alpha > 0 \\
\hline
  \small{\text{Probability density}} & \frac{\alpha\theta^\alpha}{x^{\alpha+1}} \\
 ~~ \small{\text{function }} f(x)& \\
\hline
  \small{\text{Distribution function}} & 1-(\theta/x)^\alpha \\
   ~~F(x) & \\
\hline
  \textit{k}^{th}~\small{\text{raw moment}} & \frac{\alpha\theta^k}{\alpha-k} \\
  ~~\mathrm{E}[X^k]  & k < \alpha \\
\hline
  \mathrm{E}[(X\wedge x)^k] & \frac{\alpha\theta^k}{\alpha-k}-\frac{k\theta^{\alpha}}{(\alpha-k)x^{\alpha-k}} \\
  & x \geq\theta \\
\hline
\end{array}
\end{matrix}
$$

**`R` Commands**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Function Name} & \text{R Command} \\
\hline
  \small{\text{Density function}} & \text{dpareto1}(x=, shape=\alpha,min=\theta) \\
\hline
  \small{\text{Distribution function}} & \text{ppareto1}(p=, shape=\alpha,min=\theta) \\
\hline
  \small{\text{Quantile function}} & \text{qpareto1}(q=, shape=\alpha,min=\theta) \\ 
\hline
  \small{\text{Random sampling function}} & \text{rpareto1}(n=, shape=\alpha,min=\theta) \\
\hline
\end{array}
\end{matrix}
$$

**Illustrative Graph**

```{r echo=SHOW_PDF}
alpha <- 3
theta <- 100
X <- seq(from = 0, to = 1000, by = 1)
plot(x=X,y=dpareto1(X,shape=alpha,min=theta),type="l",
     ylab="Probability density",col="red",main="Single Parameter Pareto Distribution")
```

</div>

### Two Parameter Distributions


<div class="tab">
  <button class="tablinks" onclick="openTab(event, '2pA')">Pareto</button>
  <button class="tablinks" onclick="openTab(event, '2pB')">Inv Pareto</button>
  <button class="tablinks" onclick="openTab(event, '2pC')">Loglogistic</button>
  <button class="tablinks" onclick="openTab(event, '2pD')">Paralogistic</button>
  <button class="tablinks" onclick="openTab(event, '2pE')">Gamma</button>
  <button class="tablinks" onclick="openTab(event, '2pF')">Inv Gamma</button>
  <button class="tablinks" onclick="openTab(event, '2pG')">Weibull</button>
  <button class="tablinks" onclick="openTab(event, '2pH')">Inv Weibull</button>
  <button class="tablinks" onclick="openTab(event, '2pI')">Uniform</button>
  <button class="tablinks" onclick="openTab(event, '2pJ')">Normal</button>
</div>

<div id="2pA" class="tabcontent">
  <span onclick="this.parentElement.style.display='none'" class="topright"><button>Hide</button></span>

<h3>Pareto</h3>

**Functions**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Name} & \text{Function} \\
\hline
  \small{\text{Parameter assumptions}} & \theta>0, \alpha>0 \\
\hline
  \small{\text{Probability density}} & \frac{\alpha\theta^\alpha}{(x+\theta)^{\alpha+1}} \\
 ~~ \small{\text{function }} f(x)& \\
\hline
  \small{\text{Distribution function}} & 1-\Big(\frac{\theta}{x+\theta}\Big)^\alpha \\
   ~~F(x) & \\
\hline
  \textit{k}^{th}~\small{\text{raw moment}} & \frac{\theta^k\Gamma(k+1)\Gamma(\alpha-k)}{\Gamma(\alpha)} \\
  ~~\mathrm{E}[X^k] & -1<k<\alpha \\
\hline
  \small{\text{Limited Expected Value:}}~\alpha\neq1 & \frac{\theta}{\alpha-1}\Big[1-\Big(\frac{\theta}{x+\theta}\Big)^{\alpha-1}\Big] \\
  ~~\mathrm{E}[X\wedge x] & \\
\hline
  \small{\text{Limited Expected Value:}}~\alpha=1 & -\theta \ln \left(\frac{\theta}{x+\theta}\right) \\
  ~~\mathrm{E}[X\wedge x] & \\
\hline
  \mathrm{E}[(X\wedge x)^k] & \frac{\theta^k\Gamma(k+1)\Gamma(\alpha-k)}{\Gamma(\alpha)}\beta(k+1,\alpha-k;\frac{x}{x+\theta})+x^k(\frac{\theta}{x+\theta})^\alpha  \\
\hline
\end{array}
\end{matrix}
$$


**`R` Commands**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Function Name} & \text{R Command} \\
\hline
  \small{\text{Density function}} & \text{dpareto2}(x=, shape=\alpha, scale=\theta) \\
\hline
  \small{\text{Distribution function}} & \text{ppareto2}(p=, shape=\alpha,scale=\theta) \\
\hline
  \small{\text{Quantile function}} & \text{qpareto2}(q=, shape=\alpha,scale=\theta) \\ 
\hline
  \small{\text{Random sampling function}} & \text{rpareto2}(n=, shape=\alpha,scale=\theta) \\
\hline
\end{array}
\end{matrix}
$$

**Illustrative Graph**

```{r echo=SHOW_PDF}
alpha <- 3
theta <- 200
X <- seq(from = 0, to = 1000, by = 1)
plot(x=X,y=dpareto2(X,shape=alpha,scale=theta),type="l",
     ylab="Probability density",col="red",main="Pareto Distribution")
```

</div>

<div id="2pB" class="tabcontent">
  <span onclick="this.parentElement.style.display='none'" class="topright"><button>Hide</button></span>

<h3>Inverse Pareto</h3>

**Functions**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Name} & \text{Function} \\
  \hline
  \small{\text{Parameter assumptions}} & \theta>0, \tau>0 \\
\hline
  \small{\text{Probability density}} & \frac{\tau\theta x^{\tau-1}}{(x+\theta)^\tau-1} \\
 ~~ \small{\text{function }} f(x)& \\
\hline
  \small{\text{Distribution function}} & \Big(\frac{x}{x+\theta}\Big)^\tau \\
   ~~F(x) & \\
\hline
  \textit{k}^{th}~\small{\text{raw moment}} & \frac{\theta^k\Gamma(\tau+k)\Gamma(1-k)}{\Gamma(\tau)} \\
  ~~\mathrm{E}[X^k]  & -\tau<k<1 \\
\hline
  \mathrm{E}[(X\wedge x)^k] & \theta^k\tau\int^{x/(x+\theta)}_0~y^{\tau+k-1}(1-y)^{-k}dy+x^k[1-\Big(\frac{x}{x+\theta}\Big)^\tau] \\
  & k>-\tau \\
\hline
\end{array}
\end{matrix}
$$
**`R` Commands**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Function Name} & \text{R Command} \\
\hline
  \small{\text{Density function}} & \text{dinvpareto}(x=, shape=\tau, scale=\theta) \\
\hline
  \small{\text{Distribution function}} & \text{pinvpareto}(p=, shape=\tau,scale=\theta) \\
\hline
  \small{\text{Quantile function}} & \text{qinvpareto}(q=, shape=\tau,scale=\theta) \\ 
\hline
  \small{\text{Random sampling function}} & \text{rinvpareto}(n=, shape=\tau,scale=\theta) \\
\hline
\end{array}
\end{matrix}
$$

**Illustrative Graph**

```{r echo=SHOW_PDF}
tau <- 5
theta <- 100
X <- seq(from=0,to=3000,by=1)
plot(x=X,y=dinvpareto(X,shape=tau,scale=theta),type="l",
     ylab="Probability density",col="red",main="Inverse Pareto Distribution")
```


</div>

<div id="2pC" class="tabcontent">
  <span onclick="this.parentElement.style.display='none'" class="topright"><button>Hide</button></span>

<h3>Loglogistic</h3>

**Functions**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Name} & \text{Function} \\
\hline
  \small{\text{Parameter assumptions}} & \theta>0, \gamma > 0, u=\frac{(x/\theta)^\gamma}{1+(x/\theta)^\gamma} \\
\hline
  \small{\text{Probability density}} & \frac{\gamma(x/\theta)^\gamma}{x[1+(x/\theta)^\gamma]^2} \\
 ~~ \small{\text{function }} f(x)& \\
\hline
  \small{\text{Distribution function}} & u \\
   ~~F(x) & \\
\hline
  \textit{k}^{th}~\small{\text{raw moment}} & \theta^k\Gamma(1+(k/\gamma))\Gamma(1-(k/\gamma)) \\
  ~~\mathrm{E}[X^k]  & -\gamma<k<\gamma \\
\hline
  \mathrm{E}[(X\wedge x)^k] & \theta^k\Gamma(1+(k/\gamma))\Gamma(1-(k/\gamma))\beta(1+(k/\gamma),1-(k/\gamma);u)+x^k(1-u) \\
  & k>-\gamma \\
\hline
\end{array}
\end{matrix}
$$

**Illustrative Graph**

```{r eval=TRUE, include=TRUE, echo=SHOW_PDF}
dloglogistic <- function(x, gamma, theta){
  p=gamma*(x/theta)^gamma/(x*(1+(x/theta)^gamma)^2)
  return(p)
}
gamma <- 2
theta <- 100
X <- seq(from = 0, to = 1000, by = 1)
plot(x=X,y=dloglogistic(X,gamma=gamma,theta=theta),type="l",col="red")
```


</div>

<div id="2pD" class="tabcontent">
  <span onclick="this.parentElement.style.display='none'" class="topright"><button>Hide</button></span>

<h3>Paralogistic</h3>

**Functions**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Name} & \text{Function} \\
\hline
  \small{\text{Parameter assumptions}} & \theta>0, \alpha>0, u=\frac{1}{1+(x/\theta)^\alpha} \\
\hline
  \small{\text{Probability density}} & \frac{\alpha^2(x/\theta)^\alpha}{x[1+(x/\theta)^\alpha]^{\alpha+1}} \\
 ~~ \small{\text{function }} f(x)& \\
\hline
  \small{\text{Distribution function}} & 1-u^\alpha \\
   ~~F(x) & \\
\hline
  \textit{k}^{th}~\small{\text{raw moment}} & \frac{\theta^k\Gamma(1+(k/\alpha))\Gamma(\alpha-(k/\alpha))}{\Gamma(\alpha)} \\
  ~~\mathrm{E}[X^k]  & -\alpha<k<\alpha^2 \\
\hline
  \mathrm{E}[(X\wedge x)^k] & \frac{\theta^k\Gamma(1+(k/\alpha))\Gamma(\alpha-(k/\alpha))}{\Gamma(\alpha)}\beta(1+(k/\alpha),\alpha-(k/\alpha);1-u)+x^ku^\alpha \\
  & k>-\alpha \\
\hline
\end{array}
\end{matrix}
$$



**`R` Commands**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Function Name} & \text{R Command} \\
\hline
  \small{\text{Density function}} & \text{dparalogis}(x=, shape=\alpha, scale=\theta) \\
\hline
  \small{\text{Distribution function}} & \text{pparalogis}(p=, shape=\alpha,scale=\theta) \\
\hline
  \small{\text{Quantile function}} & \text{qparalogis}(q=, shape=\alpha,scale=\theta) \\ 
\hline
  \small{\text{Random sampling function}} & \text{rparalogis}(n=, shape=\alpha,scale=\theta) \\
\hline
\end{array}
\end{matrix}
$$

**Illustrative Graph**

```{r echo=SHOW_PDF}
alpha <- 2
theta <- 100
X <- seq(from = 0, to = 1000, by = 1)
plot(x=X,y=dparalogis(X,shape=alpha,scale=theta),type="l",
     ylab="Probability density",col="red",main="Paralogistic Distribution")
```

</div>

<div id="2pE" class="tabcontent">
  <span onclick="this.parentElement.style.display='none'" class="topright"><button>Hide</button></span>

<h3>Gamma</h3>

**Functions**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Name} & \text{Function} \\
\hline
  \small{\text{Parameter assumptions}} & \theta>0,~\alpha>0 \\
\hline
  \small{\text{Probability density}} & \frac{1}{\theta^{\alpha}\Gamma(\alpha)}x^{\alpha-1}e^{-x/\theta} \\
 ~~ \small{\text{function }} f(x)& \\
\hline
  \small{\text{Distribution function}} & \Gamma(\alpha;\frac{x}{\theta}) \\
   ~~F(x) & \\
\hline
  \textit{k}^{th}~\small{\text{raw moment}}  & \theta^k\frac{\Gamma(\alpha+k)}{\Gamma(\alpha)} \\
  ~~\mathrm{E}[X^k]  & k>-\alpha \\
\hline
   & \frac{\theta^k\Gamma(k+\alpha)}{\Gamma(\alpha)}\Gamma(k+\alpha; x/\theta)+x^k[1-\Gamma(\alpha; x/\theta)]  \\
   ~~\mathrm{E}[X\wedge x]^k & k > -\alpha \\
\hline
\end{array}
\end{matrix}
$$


**`R` Commands**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \small{\text{Density function}} & \text{dgamma}(x=, shape=\alpha, scale=\theta) \\
\hline
  \small{\text{Distribution function}} & \text{pgamma}(p=, shape=\alpha,scale=\theta) \\
\hline
  \small{\text{Quantile function}} & \text{qgamma}(q=, shape=\alpha,scale=\theta) \\ 
\hline
  \small{\text{Random sampling function}} & \text{rgamma}(n=, shape=\alpha,scale=\theta) \\
\hline
\end{array}
\end{matrix}
$$

**Illustrative Graph**

```{r echo=SHOW_PDF}
alpha <- 2
theta <- 50
X <- seq(from = 0, to = 1000, by = 1)
plot(x=X,y=dgamma(X,shape=alpha,scale=theta),type="l",
     ylab="Probability density",col="red",main="Gamma Distribution")
```

</div>

<div id="2pF" class="tabcontent">
  <span onclick="this.parentElement.style.display='none'" class="topright"><button>Hide</button></span>

<h3>Inverse Gamma</h3>

**Functions**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Name} & \text{Function} \\
\hline
  \small{\text{Probability density}} & \frac{(\theta/x)^\alpha e^{-\theta/x}}{x\Gamma(\alpha)} \\
 ~~ \small{\text{function }} f(x)& \\
\hline
  \small{\text{Distribution function}} & 1-\Gamma(\alpha;\theta/x) \\
   ~~F(x) & \\
\hline
  \textit{k}^{th}~\small{\text{raw moment}} & \frac{\theta^k\Gamma(\alpha-k)}{\Gamma(\alpha)} \\
  ~~\mathrm{E}[X^k]  & k<\alpha \\
\hline
  \mathrm{E}[(X\wedge x)^k] & \frac{\theta^k\Gamma(\alpha-k)}{\Gamma(\alpha)}[1-\Gamma(\alpha-k;\theta/x)]+x^k\Gamma(\alpha;\theta/x) \\
  &  \\
\hline
\end{array}
\end{matrix}
$$


**`R` Commands**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Function Name} & \text{R Command} \\
\hline
  \small{\text{Density function}} & \text{dinvgamma}(x=, shape=\alpha, scale=\theta) \\
\hline
  \small{\text{Distribution function}} & \text{pinvgamma}(p=, shape=\alpha,scale=\theta) \\
\hline
  \small{\text{Quantile function}} & \text{qinvgamma}(q=, shape=\alpha,scale=\theta) \\ 
\hline
  \small{\text{Random sampling function}} & \text{rinvgamma}(n=, shape=\alpha,scale=\theta) \\
\hline
\end{array}
\end{matrix}
$$

**Illustrative Graph**

```{r echo=SHOW_PDF}
alpha <- 3
theta <- 100
X <- seq(from=0,to=400,by=1)
plot(x=X,y=dinvgamma(X,shape=alpha,scale=theta),type="l",
     ylab="Probability density",col="red",main="Inverse Gamma Distribution")
```

</div>

<div id="2pG" class="tabcontent">
  <span onclick="this.parentElement.style.display='none'" class="topright"><button>Hide</button></span>

<h3>Weibull</h3>


**Functions**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Name} & \text{Function} \\
\hline
  \small{\text{Parameter assumptions}} & \theta>0,\alpha>0 \\
\hline\
  \small{\text{Probability density}} & \frac{\alpha \Big(\frac{x}{\theta}\Big)^\alpha \exp\Big(-\Big(\frac{x}{\theta}\Big)^\alpha\Big)}{x} \\
 ~~ \small{\text{function }} f(x)& \\
\hline
  \small{\text{Distribution function}} & 1-\exp\Big(-\Big(\frac{x}{\theta}\Big)^\alpha\Big) \\
   ~~F(x) & \\
\hline
  \textit{k}^{th}~\small{\text{raw moment}} & \theta^k \Gamma(1 + \frac{k}{\alpha}) \\
  ~~\mathrm{E}[X^k]  & k>-\alpha \\
\hline
  \mathrm{E}[(X\wedge x)^k] & \theta^k\Gamma(1+\frac{k}{\alpha})\Gamma\Big[1+\frac{k}{\alpha};\Big(\frac{x}{\theta}\Big)^\alpha\Big]+x^k\exp\Big(-\Big(\frac{x}{\theta}\Big)^\alpha\Big)  \\
   & k>-\alpha \\
\hline
\end{array}
\end{matrix}
$$


**`R` Commands**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Function Name} & \text{R Command} \\
\hline
  \small{\text{Density function}} & \text{dweibull}(x=, shape=\alpha, scale=\theta) \\
\hline
  \small{\text{Distribution function}} & \text{pweibull}(p=, shape=\alpha,scale=\theta) \\
\hline
  \small{\text{Quantile function}} & \text{qweibull}(q=, shape=\alpha,scale=\theta) \\ 
\hline
  \small{\text{Random sampling function}} & \text{rweibull}(n=, shape=\alpha,scale=\theta) \\
\hline
\end{array}
\end{matrix}
$$

**Illustrative Graph**

```{r echo=SHOW_PDF}
alpha <- 2
theta <- 100
X <- seq(from = 0, to = 1000, by = 1)
plot(x=X,y=dweibull(X,shape=alpha, scale=theta),type="l",
     ylab="Probability density",col="red",main="Weibull Distribution")
```

</div>

<div id="2pH" class="tabcontent">
  <span onclick="this.parentElement.style.display='none'" class="topright"><button>Hide</button></span>

<h3>Inverse Weibull</h3>

**Functions**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Name} & \text{Function} \\
\hline
  \small{\text{Parameter assumptions}} & \theta>0,\tau>0 \\
\hline\
  \small{\text{Probability density}} & \frac{\tau(\theta/x)^\tau \exp\Big(-\Big(\frac{\theta}{x}\Big)^\tau\Big)}{x} \\
 ~~ \small{\text{function }} f(x)& \\
\hline
  \small{\text{Distribution function}} & \exp\Big(-\Big(\frac{\theta}{x}\Big)^\tau\Big) \\
   ~~F(x) & \\
\hline
 \textit{k}^{th}~\small{\text{raw moment}} & \theta^k\Gamma(1-(k/\tau)) \\
  ~~\mathrm{E}[X^k]  & k<\tau \\
\hline
  \mathrm{E}[(X\wedge x)^k] & \theta^k\Gamma(1-(k/\tau))[1-\Gamma(1-(k/\tau);(\theta/x)^\tau)]+x^k[1-e^{-(\theta/x)^\tau}] \\
  &  \\
\hline
\end{array}
\end{matrix}
$$

**`R` Commands**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Function Name} & \text{R Command} \\
\hline
  \small{\text{Density function}} & \text{dinvweibull}(x=, shape=\tau, scale=\theta) \\
\hline
  \small{\text{Distribution function}} & \text{pinvweibull}(p=, shape=\tau,scale=\theta) \\
\hline
  \small{\text{Quantile function}} & \text{qinvweibull}(q=, shape=\tau,scale=\theta) \\ 
\hline
  \small{\text{Random sampling function}} & \text{rinvweibull}(n=, shape=\tau,scale=\theta) \\
\hline
\end{array}
\end{matrix}
$$

**Illustrative Graph**

```{r echo=SHOW_PDF}
tau <- 5
theta <- 100
X <- seq(from = 0, to = 1000, by = 1)
plot(x=X,y=dinvweibull(X,shape=tau,scale=theta),type="l",
     ylab="Probability density",col="red",main="Inverse Weibull Distribution")
```

</div>

<div id="2pI" class="tabcontent">
  <span onclick="this.parentElement.style.display='none'" class="topright"><button>Hide</button></span>

<h3>Uniform</h3>

**Functions**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Name} & \text{Function} \\
\hline
  \small{\text{Parameter assumptions}} & -\infty<\alpha<\beta<\infty \\
\hline
  \small{\text{Probability density}} & \frac{1}{\beta-\alpha} \\
  \text{f(x)} & \\
\hline
  \small{\text{Distribution function}} & \frac{x-\alpha}{\beta-\alpha} \\
 ~~F(x) & \\
\hline
  \text{Mean} & \frac{\beta+\alpha}{2} \\
  \text{E[X]} & \\
\hline
  \text{Variance} & \frac{(\beta-\alpha)^2}{12} \\
  E[(X-\mu)^2] & \\
\hline
  \mathrm{E}[(X-\mu)^k] & \mu_k=0~~~\text{for odd }\textit{k} \\
   & \mu_k=\frac{(\beta-\alpha)^k}{2^k (k+1)}~~~\text{for even }\textit{k} \\
\hline
\end{array}
\end{matrix}
$$

**`R` Commands**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Function Name} & \text{R Command} \\
\hline
  \small{\text{Density function}} & \text{dunif}(x=, min=a, max=b) \\
\hline
  \small{\text{Distribution function}} & \text{punif}(p=, min=a, max=b) \\
\hline
  \small{\text{Quantile function}} & \text{qunif}(q=, min=a, max=b) \\ 
\hline
  \small{\text{Random sampling function}} & \text{runif}(n=, min=a, max=b) \\
\hline
\end{array}
\end{matrix}
$$

**Illustrative Graph**

```{r echo=SHOW_PDF}
alpha <- 50
beta <- 100
X <- seq(alpha,beta,1)
plot(x=X,y=dunif(X,alpha,beta),type="l",
     ylab="Probability density",col="red",main="Continuous Uniform Distribution")
```


</div>


<div id="2pJ" class="tabcontent">
  <span onclick="this.parentElement.style.display='none'" class="topright"><button>Hide</button></span>

<h3>Normal</h3>

**Functions**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Name} & \text{Function} \\
\hline
  \small{\text{Parameter assumptions}} & -\infty<\mu<\infty,~\sigma>0 \\
\hline
  \small{\text{Probability density}} & \frac{1}{\sqrt{2\pi}\sigma} \exp\left( -\frac{(x-\mu)^2}{2\sigma^2}\right) \\
  \text{f(x)} & \\
\hline
  \small{\text{Distribution function}} & \Phi\left(\frac{x-\mu}{\sigma}\right) \\
 ~~F(x) & \\
\hline
  \text{Mean} & \mu \\
  \text{E[X]} & \\
\hline
  \text{Variance} & \sigma^2 \\
  E[(X-\mu)^2] & \\
\hline
  \mathrm{E}[(x-\mu)^k] & \mu_k=0~~~\text{for even k} \\
   & \mu_k=\frac{k!\sigma^2}{(\frac{k}{2})! 2^{k/2}}~~~\text{for odd k} \\
\hline
\end{array}
\end{matrix}
$$



**`R` Commands**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Function Name} & \text{R Command} \\
\hline
  \small{\text{Density function}} & \text{dnorm}(x=, mean=\mu, sd=\sigma) \\
\hline
  \small{\text{Distribution function}} & \text{pnorm}(p=, mean=\mu, sd=\sigma) \\
\hline
  \small{\text{Quantile function}} & \text{qnorm}(q=, mean=\mu, sd=\sigma) \\ 
\hline
  \small{\text{Random sampling function}} & \text{rnorm}(n=, mean=\mu, sd=\sigma) \\
\hline
\end{array}
\end{matrix}
$$

**Illustrative Graph**

```{r echo=SHOW_PDF}
mu <- 100
sigma <- 10
X <- seq(from=0,to=200,by=1)
plot(x=X,y=dnorm(X,mean=mu,sd=sigma),type="l",
     ylab="Probability density",col="red",main="Normal Distribution")
```

</div>

<div id="2pK" class="tabcontent">
  <span onclick="this.parentElement.style.display='none'" class="topright"><button>Hide</button></span>

<h3>Cauchy</h3>

**Functions**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Name} & \text{Function} \\
\hline
  \small{\text{Parameter assumptions}} & -\infty <\alpha <\infty, \beta>0 \\
\hline
  \small{\text{Probability density}} & \frac{1}{\pi\beta}[1+\left( \frac{x-\alpha}{\beta}\right)^2]^{-1} \\
 ~~ \small{\text{function }} f(x)& \\
\hline
\end{array}
\end{matrix}
$$


**`R` Commands**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Function Name} & \text{R Command} \\
\hline
  \small{\text{Density function}} & \text{dcauchy}(x=, location=\alpha, scale=\beta) \\
\hline
  \small{\text{Distribution function}} & \text{pcauchy}(p=, location=\alpha, scale=\beta) \\
\hline
  \small{\text{Quantile function}} & \text{qcauchy}(q=, location=\alpha, scale=\beta) \\ 
\hline
  \small{\text{Random sampling function}} & \text{rcauchy}(n=, location=\alpha, scale=\beta) \\
\hline
\end{array}
\end{matrix}
$$

**Illustrative Graph**

```{r echo=SHOW_PDF}
alpha <- 50
beta <- 100
X <- seq(from = 0, to = 1000, by = 1)
plot(x=X,y=dcauchy(X,location=alpha,scale=beta),
     type="l",ylab="Probability density",col="red",main="Cauchy Distribution")
```

</div>


### Three Parameter Distributions


<div class="tab">
  <button class="tablinks" onclick="openTab(event, '3pA')">Generalized Pareto</button>
  <button class="tablinks" onclick="openTab(event, '3pB')">Burr</button>
  <button class="tablinks" onclick="openTab(event, '3pC')">Inv Burr</button>
</div>

<div id="3pA" class="tabcontent">
  <span onclick="this.parentElement.style.display='none'" class="topright"><button>Hide</button></span>

<h3>Generalized Pareto</h3>

**Functions**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Name} & \text{Function} \\
\hline
  \small{\text{Parameter assumptions}} & \theta>0, \alpha>0, \tau>0, u=\frac{x}{x+\theta} \\
\hline
  \small{\text{Probability density}} & \frac{\Gamma(\alpha+\tau)}{\Gamma(\alpha)\Gamma(\tau)}\frac{\theta^\alpha x^{\tau-1}}{(x+\theta)^{\alpha+\tau}} \\
 ~~ \small{\text{function }} f(x)& \\
\hline
  \small{\text{Distribution function}} & \beta(\tau,\alpha;u) \\
   ~~F(x) & \\
\hline
  \textit{k}^{th}~\small{\text{raw moment}} & \frac{\theta^k\Gamma(\tau+1)\Gamma(\alpha-k)}{\Gamma(\alpha)\Gamma(\tau)} \\
  ~~~~\mathrm{E}[X^k]  & -\tau<k<\alpha \\
\hline
  \mathrm{E}[(X\wedge x)^k] & \frac{\theta^k\Gamma(\tau+k)\Gamma(\alpha-k)}{\Gamma(\alpha)\Gamma(\tau)}\beta(\tau+k,\alpha-k;u)+x^k[1-\beta(\tau,\alpha;u)]  \\
  & k>-\tau \\
\hline
\end{array}
\end{matrix}
$$


**`R` Commands**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Function Name} & \text{R Command} \\
\hline
  \small{\text{Density function}} & \text{dgenpareto}(x=, shape1=\alpha, shape2=\tau, scale=\theta) \\
\hline
  \small{\text{Distribution function}} & \text{pgenpareto}(q=, shape1=\alpha, shape2=\tau, scale=\theta) \\
\hline
  \small{\text{Quantile function}} & \text{qgenpareto}(p=, shape1=\alpha, shape2=\tau, scale=\theta) \\ 
\hline
  \small{\text{Random sampling function}} & \text{rgenpareto}(r=, shape1=\alpha, shape2=\tau, scale=\theta) \\
\hline
\end{array}
\end{matrix}
$$

**Illustrative Graph**

```{r echo=SHOW_PDF}
alpha <- 3
tau <- 5
theta <- 100
X <- seq(from = 0, to = 1000, by = 1)
plot(x=X,y=dgenpareto(X,shape1=alpha,shape2=tau,scale=theta),type="l",
     ylab="Probability density",col="red",main="Generalized Pareto Distribution")
```

</div>

<div id="3pB" class="tabcontent">
  <span onclick="this.parentElement.style.display='none'" class="topright"><button>Hide</button></span>

<h3>Burr</h3>

**Functions**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Name} & \text{Function} \\
\hline
  \small{\text{Parameter assumptions}} & \theta>0, \alpha>0, \gamma>0, u=\frac{1}{1+(x/\theta)^\gamma} \\
\hline
  \small{\text{Probability density}} & \frac{\alpha\gamma(x/\theta)^\gamma}{x[1+(x/\theta)^\gamma]^{\alpha+1}} \\
 ~~ \small{\text{function }} f(x)& \\
\hline
  \small{\text{Distribution function}} & 1-u^\alpha \\
   ~~F(x) & \\
\hline
  \textit{k}^{th}~\small{\text{raw moment}} & \frac{\theta^k\Gamma(1+(k/\gamma))\Gamma(\alpha-(k/\gamma))}{\Gamma(\alpha)} \\
  ~~~~\mathrm{E}[X^k]  & -\gamma<k<\alpha\gamma \\
\hline
  \mathrm{E}[(X\wedge x)^k] & \frac{\theta^k\Gamma(1+(k/\gamma))\Gamma(\alpha-(k/\gamma))}{\Gamma(\alpha)}\beta(1+(k/\gamma),\alpha-(k/\gamma);1-u)+x^ku^\alpha  \\
  & k>-\gamma \\
\hline
\end{array}
\end{matrix}
$$


**`R` Commands**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Function Name} & \text{R Command} \\
\hline
  \small{\text{Density function}} & \text{dburr}(x=, shape1=\alpha, shape2=\gamma, scale=\theta) \\
\hline
  \small{\text{Distribution function}} & \text{pburr}(p=, shape1=\alpha, shape2=\gamma, scale=\theta) \\
\hline
  \small{\text{Quantile function}} & \text{qburr}(q=, shape1=\alpha, shape2=\gamma, scale=\theta) \\ 
\hline
  \small{\text{Random sampling function}} & \text{rburr}(n=, shape1=\alpha, shape2=\gamma, scale=\theta) \\
\hline
\end{array}
\end{matrix}
$$

**Illustrative Graph**

```{r echo=SHOW_PDF}
alpha <- 2
gamma <- 3
theta <- 100
X <- seq(from = 0, to = 1000, by = 1)
plot(x=X,y=dburr(X,shape1=alpha,shape2=gamma,scale=theta),type="l",
     ylab="Probability density",col="red",main="Burr Distribution")
```

</div>

<div id="3pC" class="tabcontent">
  <span onclick="this.parentElement.style.display='none'" class="topright"><button>Hide</button></span>

<h3>Inverse Burr</h3>

**Functions**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Name} & \text{Function} \\
\hline
  \small{\text{Parameter assumptions}} & \theta>0, \tau>0, \gamma>0, u=\frac{(x/\theta)^\gamma}{1+(x/\theta)^\gamma} \\
\hline
  \small{\text{Probability density}} & \frac{\tau\gamma(x/\theta)^{\tau \gamma}}{x[1+(x/\theta)^\gamma]^{\tau+1}} \\
 ~~ \small{\text{function }} f(x)& \\
\hline
  \small{\text{Distribution function}} & u^\tau \\
   ~~F(x) & \\
\hline
  \textit{k}^{th}~\small{\text{raw moment}} & \frac{\theta^k\Gamma(\tau+(k/\gamma))\Gamma(1-(k/\gamma))}{\Gamma(\tau)} \\
  ~~~~\mathrm{E}[X^k]  & -\tau\gamma<k<\gamma \\
\hline
  \mathrm{E}[(X\wedge x)^k] & \frac{\theta^k\Gamma(\tau+(k/\gamma))\Gamma(1-(k/\gamma))}{\Gamma(\tau)}\beta(\tau+(k/\gamma),1-(k/\gamma);u)+x^k[1-u^\tau]  \\
  & k>-\tau\gamma \\
\hline
\end{array}
\end{matrix}
$$

**`R` Commands**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Function Name} & \text{R Command} \\
\hline
  \small{\text{Density function}} & \text{dinvburr}(x=, shape1=\tau, shape2=\gamma, scale=\theta) \\
\hline
  \small{\text{Distribution function}} & \text{pinvburr}(p=, shape1=\tau, shape2=\gamma, scale=\theta) \\
\hline
  \small{\text{Quantile function}} & \text{qinvburr}(q=, shape1=\tau, shape2=\gamma, scale=\theta) \\ 
\hline
  \small{\text{Random sampling function}} & \text{rinvburr}(n=, shape1=\tau, shape2=\gamma, scale=\theta) \\
\hline
\end{array}
\end{matrix}
$$

**Illustrative Graph**

```{r echo=SHOW_PDF}
tau <- 2
gamma <- 3
theta <- 100
X <- seq(from = 0, to = 1000, by = 1)
plot(x=X,y=dinvburr(X,shape1=tau,shape2=gamma,scale=theta),type="l",
     ylab="Probability density",col="red",main="Inverse Burr Distribution")
```

</div>



### Four Parameter Distribution


<div class="tab">
  <button class="tablinks" onclick="openTab(event, '4pA')">GB2</button>
</div>

<div id="4pA" class="tabcontent">
  <span onclick="this.parentElement.style.display='none'" class="topright"><button>Hide</button></span>

<h3>Generalized Beta of the Second Kind (GB2)</h3>

**Functions**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Name} & \text{Function} \\
\hline
  \small{\text{Parameter assumptions}} & \theta>0, \alpha_1>0, \alpha_2>0, \sigma>0  \\
\hline
  \small{\text{Probability density}} & \frac{(x/\theta)^{\alpha_2/\sigma}}{x \sigma~\mathrm{B}\left( \alpha_1,\alpha_2\right)\left\lbrack 1 + \left( x/\theta \right)^{1/\sigma} \right\rbrack^{\alpha_1 + \alpha_2}}  \\
    ~~ \small{\text{function }} f(x) & \\
\hline
  \textit{k}^{th}~\small{\text{raw moment}} & \frac{\theta^{k}~\mathrm{B}\left( \alpha_1 +k \sigma,\alpha_2 - k \sigma \right)}{\mathrm{B}\left( \alpha_1,\alpha_2 \right)} \\
  ~~~~\mathrm{E}[X^k]  & \textit{k}>0 \\
\hline
\end{array}
\end{matrix}
$$


**`R` Commands**

Please see the [R Codes for Loss Data Analytics](https://ewfrees.github.io/LDARcode/index.html) site for information about this distribution.



### Other Distributions

</center>

<div class="tab">
  <button class="tablinks" onclick="openTab(event, 'odA')">Lognormal</button>
  <button class="tablinks" onclick="openTab(event, 'odB')">Inv Gaussian</button>
  <!-- <button class="tablinks" onclick="openTab(event, 'odC')">Single-Parameter Pareto</button> -->
</div>


<div id="odA" class="tabcontent">
  <span onclick="this.parentElement.style.display='none'" class="topright"><button>Hide</button></span>

<h3>Lognormal</h3>

**Functions**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Name} & \text{Function} \\
\hline
  \small{\text{Parameter assumptions}} & -\infty <\mu <\infty, \sigma>0 \\
\hline
  \small{\text{Probability density}} & \frac{1}{x\sqrt{2\pi}\sigma} \exp\left( -\frac{(\ln x-\mu)^2}{2\sigma^2}\right) \\
 ~~ \small{\text{function }} f(x)& \\
\hline
  \small{\text{Distribution function}} & \Phi\left(\frac{\ln (x)-\mu}{\sigma}\right) \\
   ~~F(x) & \\
\hline
  \textit{k}^{th}~\small{\text{raw moment}} & \exp(k\mu+\frac{k^2\sigma^2}{2}) \\
  ~~\mathrm{E}[X^k]  &  \\
\hline
  \small{\text{Limited Expected Value}} & \exp\Big(k\mu+\frac{k^2\sigma^2}{2}\Big)\Phi\Big(\frac{\ln (x)-\mu-k\sigma^2}{\sigma}\Big)+x^k\Big[1-\Phi\Big(\frac{\ln (x)-\mu}{\sigma}\Big)\Big]  \\
  ~~\mathrm{E}[X\wedge x] & \\
\hline
\end{array}
\end{matrix}
$$

**Illustrative Graph**

```{r eval=TRUE, include=TRUE, echo=SHOW_PDF}
dlognorm <- function(x,mu,sigma){
  p=(1/(x*sigma*sqrt(2*pi)))*exp(-((log(x)-mu)/sigma)^2)
  return(p)
}
mu <- 20
sigma <- 12
X <- seq(from = 0, to = 1000, by = 1)
plot(x=X,y=dlognorm(X,mu=mu,sigma=sigma),type="l",col="red")
```


</p>
</div>

<div id="odB" class="tabcontent">
  <span onclick="this.parentElement.style.display='none'" class="topright"><button>Hide</button></span>

<h3>Inverse Gaussian</h3>

**Functions**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Name} & \text{Function} \\
\hline
  \small{\text{Parameter assumptions}} & \theta>0, \mu>0, z=\frac{x-\mu}{\mu}~,~y=\frac{x+\mu}{\mu} \\
\hline
  \small{\text{Probability density}} & \Big(\frac{\theta}{2\pi x^3}\Big)^{1/2}\exp\Big(\frac{-\theta z^2}{2x}\Big) \\
 ~~ \small{\text{function }} f(x)& \\
\hline
  \small{\text{Distribution function}} & \Phi\Big[z\Big(\frac{\theta}{x}\Big)^{1/2}\Big]+\exp\Big(\frac{2\theta}{\mu}\Big)\Phi\Big[-y\Big(\frac{\theta}{x}\Big)^{1/2}\Big] \\
   ~~F(x) & \\
\hline
  \text{Mean} & \mu \\
  \mathrm{E}[X]  & \\
\hline
  \mathrm{Var[X]} & \frac{\mu^3}{\theta}\\
\hline
  \mathrm{E}[(X\wedge x)^k] & x-\mu x\Phi\Big[z\Big(\frac{\theta}{x}\Big)^{1/2}\Big]-(\mu y)\exp\Big(\frac{2\theta}{\mu}\Big)\Phi\Big[-y\Big(\frac{\theta}{x}\Big)^{1/2}\Big] \\
  &  \\
\hline
\end{array}
\end{matrix}
$$


**`R` Commands**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Function Name} & \text{R Command} \\
\hline
  \small{\text{Density function}} & \text{dinvgauss}(x=, mean=\mu,dispersion=\theta) \\
\hline
  \small{\text{Distribution function}} & \text{pinvgauss}(p=, mean=\mu,dispersion=\theta) \\
\hline
  \small{\text{Quantile function}} & \text{qinvgauss}(q=, mean=\mu,dispersion=\theta) \\ 
\hline
  \small{\text{Random sampling function}} & \text{rinvgauss}(n=, mean=\mu,dispersion=\theta) \\
\hline
\end{array}
\end{matrix}
$$

**Illustrative Graph**

```{r echo=SHOW_PDF}
mu <- 100
theta <- 1000
X <- seq(from=0,to=100,by=1)
plot(x=X,y=dinvgauss(X,mean=mu,dispersion=theta),type="l",
     ylab="Probability density",col="red",main="Inverse Gaussian Distribution")
```


</div>


### Distributions with Finite Support

</center>

<div class="tab">
  <button class="tablinks" onclick="openTab(event, 'fsA')">Beta</button>
  <button class="tablinks" onclick="openTab(event, 'fsB')">Generalized Beta</button>
</div>


<div id="fsA" class="tabcontent">
  <span onclick="this.parentElement.style.display='none'" class="topright"><button>Hide</button></span>

<h3>Beta</h3>

**Functions**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Name} & \text{Function} \\
\hline
  \small{\text{Parameter assumptions}} & \theta>0, ~a>0,~b>0, u=\frac{x}{\theta},~0<x<\theta \\
\hline
  \small{\text{Probability density}} & \frac{\Gamma(a+b)}{\Gamma(a)\Gamma(b)} u^a(1-u)^{b-1}\frac{1}{x} \\
 ~~ \small{\text{function }} f(x)& \\
\hline
  \small{\text{Distribution function}} & \beta(a,b;u) \\
   ~~F(x) & \\
\hline
  \textit{k}^{th}~\small{\text{raw moment}} & \frac{\theta^k 
  \Gamma(a+b)\Gamma(a+k)}{\Gamma(a)\Gamma(a+b+k)} \\
  ~~\mathrm{E}[X^k]  & k>-a \\
\hline
  & \frac{\theta^k a(a+1)\cdots(a+k-1)}{(a+b)(a+b+1)\cdots(a+b+k-1)}\beta(a+k,b;u)+x^k[1-\beta(a,b;u)]  \\
   ~~\mathrm{E}[X\wedge x]^k &  \\
\hline
\end{array}
\end{matrix}
$$

**`R` Commands**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Function Name} & \text{R Command} \\
\hline
  \small{\text{Density function}} & \text{dbeta}(x=, shape1=a,shape2=b,ncp=\theta) \\
\hline
  \small{\text{Distribution function}} & \text{pbeta}(p=, shape1=a,shape2=b,ncp=\theta) \\
\hline
  \small{\text{Quantile function}} & \text{qbeta}(q=, shape1=a,shape2=b,ncp=\theta) \\ 
\hline
  \small{\text{Random sampling function}} & \text{rbeta}(n=, shape1=a,shape2=b,ncp=\theta) \\
\hline
\end{array}
\end{matrix}
$$


```{r echo=SHOW_PDF}
a <- 2
b <- 4
theta <- 1
X <- seq(from=0,to=1,by=.0001)
plot(x=X,y=dbeta(X,shape1=a,shape2=b,ncp=theta),type="l",
     ylab="Probability density",col="red",main="Beta Distribution")
```


</div>

<div id="fsB" class="tabcontent">
  <span onclick="this.parentElement.style.display='none'" class="topright"><button>Hide</button></span>

<h3>Generalized Beta</h3>

**Functions**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Name} & \text{Function} \\
\hline
  \small{\text{Parameter assumptions}} & \theta>0, a>0, b>0, \tau>0, 0<x<\theta~,~u=(x/\theta)^\tau \\
\hline
  \small{\text{Probability density}} & \frac{\Gamma(a+b)}{\Gamma(a)\Gamma(b)}u^\alpha(1-u)^{b-1}\frac{\tau}{x} \\
 ~~ \small{\text{function }} f(x)& \\
\hline
  \small{\text{Distribution function}} & \beta(a,b;u) \\
   ~~F(x) & \\
\hline
  \textit{k}^{th}~\small{\text{raw moment}} & \frac{\theta^k\Gamma(a+b)\Gamma(a+(k/\tau))}{\Gamma(a)\Gamma(a+b+(k/\tau))} \\
   ~~\mathrm{E}[X^k] & k>-\alpha\tau \\
\hline
  \mathrm{E}[(X\wedge x)^k] & \frac{\theta^k\Gamma(a+b)\Gamma(a+(k/\tau))}{\Gamma(a)\Gamma(a+b+(k/\tau))}\beta(a+(k/\tau),b;u)+x^k[1-\beta(a,b;u)] \\
\hline
\end{array}
\end{matrix}
$$

**`R` Commmands**

$$
\begin{matrix}
\begin{array}{l|c}
\hline
  \text{Function Name} & \text{R Command} \\
\hline
  \small{\text{Density function}} & \text{dgenbeta}(x=, shape1=a,shape2=b,shape3=\tau,scale=\theta) \\
\hline
  \small{\text{Distribution function}} & \text{pgenbeta}(p=, shape1=a,shape2=b,shape3=\tau,scale=\theta) \\
\hline
  \small{\text{Quantile function}} & \text{qgenbeta}(q=, shape1=a,shape2=b,shape3=\tau,scale=\theta) \\ 
\hline
  \small{\text{Random sampling function}} & \text{rgenbeta}(n=, shape1=a,shape2=b,shape3=\tau,scale=\theta) \\
\hline
\end{array}
\end{matrix}
$$

**Illustrative Graph**

```{r echo=SHOW_PDF}
a <- 3
b <- 5
tau <- 2
theta <- 1000
X <- seq(from = 0, to = 1000, by = 1)
plot(x=X,y=dgenbeta(X,shape1=a,shape2=b,shape3=tau,scale=theta),type="l",
     ylab="Probability density",col="red",main="Generalized Beta Distribution")
```

</div>

