<!-- ## Importance Sampling -->

Section \@ref(S:SimulationFundamentals) introduced Monte Carlo techniques using the inversion technique : to generate a random variable $X$ with distribution $F$, apply $F^{-1}$ to calls of a random generator (uniform on the unit interval). What if we what to draw according to $X$, conditional on $X\in[a,b]$ ?

One can use an `r Gloss('accept-reject mechanism')` : draw $x$ from distribution $F$

- if $x\in[a,b]$ : keep it ("*accept*")
- if $x\notin[a,b]$ : draw another one ("*reject*")

Observe that from $n$ values initially generated, we keep here only $[F(b)-F(a)]\cdot n$ draws, on average.



**Example `r chapnum`.4.1. Draws from a Normal Distribution.** Suppose that we draw from a normal distribution with mean 2.5 and variance 1, $N(2.5,1)$, but are only interested in draws greater that $a \ge 2$ and less than $b \le 4$. That is, we can only use $F(4)-F(2) = \Phi(4-2.5)-\Phi(2-2.5)$ = `r round(pnorm(4,2.5,1),digits=4)` - `r round(pnorm(2,2.5,1),digits=4)` = `r round(pnorm(4,2.5,1)-pnorm(2,2.5,1),digits=4)` proportion of the draws. Figure \@ref(fig:sampleani1) demonstrates that some draws lie with the interval $(2,4)$ and some are outside.

`r HideRCode('ImportanceSampling.1','Show R Code for Accept-Reject Mechanism')`

```{r eval = ANIMATION, comment = "", fig.align='center', out.width='50%', fig.asp=.75, echo=SHOW_PDF, animation.hook=ANIMATIONHOOK}
mu = 2.5
sigma = 1
a = 2
b = 4
Fa = pnorm(a,mu,sigma)
Fb = pnorm(b,mu,sigma)
pic_ani = function(){
  u=seq(0,5,by=.01)
  plot(u,pnorm(u,mu,sigma),col="white",ylab="",xlab="")
  rect(-1,-1,6,2,col=rgb(1,0,0,.2),border=NA)
  rect(a,Fa,b,Fb,col="white",border=NA)
  lines(u,pnorm(u,mu,sigma),lwd=2)
  abline(v=c(a,b),lty=2,col="red")
  ru <- runif(1)
  clr <- "red"
  if((qnorm(ru,mu,sigma)>=a)&(qnorm(ru,mu,sigma)<=b)) clr <- "blue"
  segments(-1,ru,qnorm(ru,mu,sigma),ru,col=clr,lwd=2)
  arrows(qnorm(ru,mu,sigma),ru,qnorm(ru,mu,sigma),0,col=clr,lwd=2,length = .1)
}

```

</div>

```{r sampleani1, animation.hook=ANIMATIONHOOK, fig.cap='Animated Demonstration of Draws In and Outside of (2,4)', eval = ANIMATION, comment = "", fig.align='center', out.width='50%', fig.asp=.75, echo=SHOW_PDF}
for (i in 1:numAnimation) {pic_ani()}
```

***

Instead, one can draw according to the conditional distribution $F^{\star}$ defined as

$$
F^{\star}(x) = \Pr(X \le x | a < X \le b) =\frac{F(x)-F(a)}{F(b)-F(a)}, \ \  \ \text{for } a < x \le b .
$$

Using the inverse transform method in Section \@ref(S:InverseTransform), we have that the draw

$$
X^\star=F^{\star-1}\left( U \right) = F^{-1}\left(F(a)+U\cdot[F(b)-F(a)]\right)
$$

has distribution $F^{\star}$. Expressed another way, define
$$
\tilde{U} = (1-U)\cdot F(a)+U\cdot F(b)
$$
and then use $F^{-1}(\tilde{U})$. With this approach, each draw counts.

This can be related to the `r Gloss('importance sampling mechanism')` : we draw more frequently in regions where we expect to have quantities that have some interest. This transform can be considered as a "a change of measure."

`r HideRCode('ImportanceSampling.2','Show R Code for Importance Sampling by the Inverse Transform Method')`

```{r eval = ANIMATION, echo=SHOW_PDF, animation.hook=ANIMATIONHOOK}
pic_ani = function(){
  u=seq(0,5,by=.01)
  plot(u,pnorm(u,mu,sigma),col="white",ylab="",xlab="")
  rect(-1,-1,6,2,col=rgb(1,0,0,.2),border=NA)
  rect(a,Fa,b,Fb,col="white",border=NA)
  lines(u,pnorm(u,mu,sigma),lwd=2)
  abline(h=pnorm(c(a,b),mu,sigma),lty=2,col="red")
  ru <- runif(1)
  rutilde <- (1-ru)*Fa+ru*Fb
  segments(-1,rutilde,qnorm(rutilde,mu,sigma),rutilde,col="blue",lwd=2)
  arrows(qnorm(rutilde,mu,sigma),rutilde,qnorm(rutilde,mu,sigma),0,col="blue",lwd=2,length = .1)
}

```

</div>

```{r sampleani_IS_2, eval = ANIMATION, fig.align='center', out.width='50%', fig.asp=.75, echo=SHOW_PDF, animation.hook=ANIMATIONHOOK}
for (i in 1:numAnimation) {pic_ani()}
```

In Example `r chapnum`.4.1., the inverse of the normal distribution is readily available (in `R`, the function is `qnorm`). However, for other applications, this is not the case. Then, one simply uses numerical methods to determine $X^\star$ as the solution of the equation $F(X^\star) =\tilde{U}$ where $\tilde{U}=(1-U)\cdot F(a)+U\cdot F(b)$. See the following illustrative code.


`r HideRCode('ImportanceSampling.3','Show R Code for Importance Sampling via Numerical Solutions')`

```{r eval = ANIMATION, comment = "", echo=SHOW_PDF, animation.hook=ANIMATIONHOOK}
pic_ani = function(){
  u=seq(0,5,by=.01)
  plot(u,pnorm(u,mu,sigma),col="white",ylab="",xlab="")
  rect(-1,-1,6,2,col=rgb(1,0,0,.2),border=NA)
  rect(2,-1,4,2,col="white",border=NA)
  lines(u,pnorm(u,mu,sigma),lty=2)
  pnormstar <- Vectorize(function(x){
    y=(pnorm(x,mu,sigma)-Fa)/(Fb-Fa)
    if(x<=a) y <- 0
    if(x>=b) y <- 1
    return(y)
    })
  qnormstar <- function(u) as.numeric(uniroot((function (x) pnormstar(x) - u), lower = 2, upper = 4)[1])
  lines(u,pnormstar(u),lwd=2)
  abline(v=c(2,4),lty=2,col="red")
  ru <- runif(1)
  segments(-1,ru,qnormstar(ru),ru,col="blue",lwd=2)
  arrows(qnormstar(ru),ru,qnormstar(ru),0,col="blue",lwd=2,length = .1)
}

```

</div>

```{r sampleani_IS_1, eval = ANIMATION, comment = "", fig.align='center', out.width='50%', fig.asp=.75, echo=SHOW_PDF, animation.hook=ANIMATIONHOOK}
for (i in 1:numAnimation) {pic_ani()}
```

