


# Appendix D: Summary of Distributions {#C:SummaryDistributions}



**User Notes**

* The `R` functions are from the packages `actuar` and `invgamma`.
* Tables appear when first loaded by the browser. To hide them, click on one of the distributions, e.g., *Poisson*, and then click on the *Hide* button.
* More information on the `R` codes is available at the [R Codes for Loss Data Analytics](https://ewfrees.github.io/LDARcode/index.html) site.



## Discrete Distributions

**Overview.** This section summarizes selected discrete probability distributions used throughout *Loss Data Analytics*. Relevant functions and `R` code are provided.


```{r child = 'Chapters/DistributionTableSections/Discrete.Rmd', eval = TRUE}
```

## Continuous Distributions


**Overview.** This section summarizes selected continuous probability distributions used throughout *Loss Data Analytics*. Relevant functions, `R` code, and illustrative graphs are provided.

```{r child = 'Chapters/DistributionTableSections/Continuous.Rmd', eval = TRUE}
```


## Limited Expected Values

**Overview.** This section summarizes limited expected values for selected continuous distributions.

```{r child = 'Chapters/DistributionTableSections/LimitedExpValue.Rmd', eval = TRUE}
```

