


# Appendix D: Summary of Distributions {#C:SummaryDistributions}


***

**This Summary is being written and is not yet complete nor edited. It is here to give you a flavor of what will be in the final version.**

***

**User Notes**

* The `R` functions are from the package `actuar`.
* Tables appear when first loaded by the browser. To hide them, click on one of the distributions, e.g., *Poisson*, and then click on the *Hide* button.



## Discrete Distributions

**Overview.** This section summarizes selected discrete probability distributions used throughout *Loss Data Analytics*. Relevant functions and `R` code are provided.


```{r child = 'Chapters/DistributionTableSections/Discrete.Rmd'}
```

## Continuous Distributions


**Overview.** This section summarizes selected continuous probability distributions used throughout *Loss Data Analytics*. Relevant functions, `R` code, and illustrative graphs are provided.

```{r child = 'Chapters/DistributionTableSections/Continuous.Rmd'}
```


## Limited Expected Values

**Overview.** This section summarizes limited expected values for selected continuous distributions.

```{r child = 'Chapters/DistributionTableSections/LimitedExpValue.Rmd'}
```

