#Introduction to Loss Data Analytics {#C:Intro}

*Chapter Preview*. This book introduces readers to methods of analyzing insurance data. Section \@ref(S:Intro) begins with a discussion of why the use of data is important in the insurance industry. Section \@ref(S:PredModApps) gives a general overview of the purposes of analyzing insurance data which is reinforced in the
Section \@ref(S:LGPIF) case study. Naturally, there is a huge gap between the broad goals summarized in the overview and a case study application; this gap is covered through the methods and techniques of data analysis covered
in the rest of the text.


## Relevance of Analytics {#S:Intro}

***
In this section, you learn how to:

- Summarize the importance of insurance to consumers and the economy
- How to describe analytics
- To identify data generating events associated with the timeline of a typical insurance contract

***


This book introduces the process of using data to make decisions in an
insurance context. It does not assume that readers are familiar with
insurance but introduces insurance concepts as needed. Insurance may not
be as entertaining as the sports industry (another industry that depends heavily on data) but it does affect the financial livelihoods of many. By almost any measure, insurance is a major economic activity. On a
global level, insurance premiums comprised about 6.2% of the world gross
domestic product (GDP) in 2014, [@III2016]. As examples, premiums
accounted for 18.9% of GDP in Taiwan (the highest in the study) and
represented 7.3% of GDP in the United States. On a personal level,
almost everyone owning a home has insurance to protect themselves in the
event of a fire, hailstorm, or some other calamitous event. Almost every
country requires insurance for those driving a car. In sum, although not
particularly entertaining, insurance plays an important
role in the economies of nations and the lives of individuals.

Insurance is a data-driven industry. Like other major corporations, insurers use data when trying to decide how much to pay employees, how many employees to retain, how to market their services, how to forecast
financial trends, and so on. These represent general areas of activities that are not
specific to the insurance industry. Although each industry retains its own
nuances, you will find that the methods and tools introduced in this text are relevant across industries.

When introducing data methods, we will focus on losses that arise from obligations in insurance contracts. This could be the amount of damage to one's apartment under a renter's insurance agreement, the amount needed to compensate someone that you hurt in a driving accident, and the like. We  call these obligations <a href="#" class="tooltip" style="color:green">*insurance claims*<span style="font-size:8pt"> An insurance claim is the compensation provided by the insurer for incurred hurt, loss, or damage that is covered by the policy.</span></a>. With this focus, we will be able to introduce generally applicable statistical tools and techniques in real-life situations.

### What is Analytics?

Insurance is a data-driven industry and analytics is a key to deriving
information from data. But what is analytics? Making data-driven
business decisions has been described as business analytics, business
intelligence, and data science. These terms, among others, are
sometimes used interchangeably and sometimes refer to distinct applications. *Business intelligence* may focus on processes of collecting data, often
through databases and data warehouses, whereas *business analytics*
utilizes tools and methods for statistical analyses of data. In contrast
to these two terms that emphasize business applications, the term *data
science* can encompass broader applications in many scientific domains.
For our purposes, we use the term <a href="#" class="tooltip" style="color:green">*analytics*<span style="font-size:8pt"> Analytics is the process of using data to make decisions.</span></a> to refer to the process of using data to make decisions. This process involves gathering data, understanding models of uncertainty, making general inferences, and communicating results.

### Short and Long-term Insurance

This text will focus on short-term insurance contracts. By short-term,
we mean contracts where the insurance coverage is typically provided for
six months or a year. If you are new to insurance, then it is probably
easiest to think about an insurance policy that covers the contents of
an apartment or house that you are renting (known as 
<a href="#" class="tooltip" style="color:green">renters insurance<span style="font-size:8pt">Renters insurance is an insurance policy that covers the contents of an apartment or house that you are renting.</span> </a>) or the contents and property of a building that is owned by you or a friend (known as 
<a href="#" class="tooltip" style="color:green">homeowners insurance<span style="font-size:8pt">Homeowners insurance is an insurance policy that covers the contents and property of a building that is owned by you or a friend.</span></a>). Another easy example is <a href="#" class="tooltip" style="color:green">automobile insurance<span style="font-size:8pt">An insurance policy that covers damage to your vehicle, damage to other vehicles in the accident, as well as medical expenses of those injured in the accident.</span></a>. In the event of an accident, this policy may cover damage to your vehicle, damage to other vehicles in the accident, as well as medical expenses of those injured in the accident.

In the US, policies such as renters and homeowners are known as
<a href="#" class="tooltip" style="color:green">property insurance<span style="font-size:8pt">Property insurance is a policy that protects the insured against loss or damage to real or personal property. The cause of loss might be  fire, lightening, business interruption, loss of rents, glass breakage, tornado, windstorm, hail, water damage, explosion, riot, civil commotion, rain, or damage from aircraft or vehicles.</span></a> whereas a policy such as auto that covers medical
damages to people is known as <a href="#" class="tooltip" style="color:green">casualty insurance<span style="font-size:8pt">Casualty insurance is a form of liability insurance providing coverage for negligent acts and omissions. Examples include workers compensation, errors and omissions, fidelity, crime, glass, boiler, and various malpractice coverages.</span></a> In the rest of the world, these are both known as
<a href="#" class="tooltip" style="color:green"span class="tooltip">nonlife<span style="font-size:8pt">Nonlife insurance is any type of insurance where payments are not based on the death (or survivorship) of a named insured. Examples include automobile, homeowners, and so on. Also known as property and casualty or general insurance.</span></a> or general insurance, to distinguish them from life insurance.

Both life and nonlife insurances are important components of the world economy. The
@III2016 estimates that direct insurance premiums in the world for
2014 was 2,654,549 for life and 2,123,699 for nonlife; these figures are
in *millions of US dollars*. As noted earlier, the total represents 6.2%
of the world GDP. Put another way, life accounts for 55.5% of insurance
premiums and 3.4% of world GDP whereas nonlife accounts for 44.5% of insurance
premiums and 2.8% of world GDP. Both life and nonlife represent
important economic activities and are worthy of study in their own
right.

Yet, life insurance considerations differ from nonlife. In life
insurance, the default is to have a multi-year contract. For example, if
a person 25 years old purchases a whole life policy that pays upon death
of the insured and that person does not die until age 100, then the
contract is in force for 75 years. We think of this as a long-term
contract.

Further, in life insurance, the benefit amount is often stipulated in
the contract provisions. In contrast, most short-term contracts provide
for compensation of insured losses which are unknown before the
accident. (There are usually limits placed on the
compensation amounts.) In a life insurance contract that stretches over many years, the
time value of money plays a prominent role. In contrast, in a short-term
nonlife contract, the random amount of compensation takes priority.

In both life and nonlife insurances, the frequency of claims is very
important. For many life insurance contracts, the insured event (such as
death) happens only once. In contrast, for nonlife insurances such as
automobile, it is common for individuals (especially young male drivers)
to get into more than one accident during a year. So, our models need to
reflect this observation; we will introduce different frequency models
than you may also see when studying life insurance.

For short-term insurance, the framework of the probabilistic model is
straightforward. We think of a one-period model (the period length,
e.g., six months, will be specified in the situation).

-   At the beginning of the period, the insured pays the insurer a known
    premium that is agreed upon by both parties to the contract.

-   At the end of the period, the insurer reimburses the insured for a
    (possibly multivariate) random loss.

This framework will be developed as we proceed but we first focus on
integrating this framework with concerns about how the data may arise.

### Insurance Processes {#S:InsProcesses}

One way to describe the data arising from operations of a company that
sells insurance products is to use a granular approach. In this
micro oriented view, we can think  about what happens to a
contract at various stages of its existence. Figure \@ref(fig:StochOperations) traces a timeline of a typical insurance contract. Throughout the life of the contract, the company regularly processes events such as premium collection and valuation,
described in Section \@ref(S:PredModApps); these are marked with an **x** on
the timeline. Non-regular and unanticipated events also occur.
To illustrate, $\mathrm{t}_2$ and $\mathrm{t}_4$ mark the
event of an insurance claim (some contracts, such as life insurance, can
have only a single claim). Times $\mathrm{t}_3$ and $\mathrm{t}_5$
mark events when a policyholder wishes to alter certain contract
features, such as the choice of a deductible or the amount of coverage.
Moreover, from a company perspective, one can even think about the
contract initiation (arrival, time $\mathrm{t}_1$) and contract
termination (departure, time $\mathrm{t}_6$) as uncertain events.

(ref:StochOperations) Timeline of a Typical Insurance Policy. Arrows mark the occurrences of random events. Each x marks the time of scheduled events that are typically non-random.

```{r StochOperations, fig.cap='(ref:StochOperations)', out.width='100%', fig.asp=.75, fig.align='center', echo=FALSE}
plot.new() 
par(mar = c(0,0,0,0), cex = 0.9)
ytop <- 3.5
ytop1 <- ytop - 0.5
plot.window(xlim = c(0,20), ylim = c(0,5))
#  Main Time Line
arrows(0.4, 1 , 19.8, 1, code = 2, lwd = 2, angle = 25, length = 0.10)
#  Contract Initiation
text(0.2, ytop, labels="Policy \nEnters\nPremium \nPaid", adj = 0, cex = 0.8)
arrows(0.8, ytop1, 0.8, 1.2, code = 2, lwd = 2, angle = 25, length = 0.10)
text(3.7, ytop, labels="Claim \nOccurs", adj = 0, cex = 0.8)
arrows(4.3, ytop1, 4.3, 1.2, code = 2, lwd = 2, angle = 25, length = 0.10)

text(9,ytop,labels="Contract \nAlteration\nIncrease \nDeductible", adj = 0, cex = 0.8)
arrows(9.1,ytop1,9.1,1.2,code=2,lwd=2,angle=25,length=0.10)

text(9.2,1.8,labels="Policy \nRenewal\nPremium \nPaid", adj = 0, cex = 0.8)
text(9,1,labels="X", adj = 0, cex = 0.8)

text(12,ytop,labels="Claim \nOccurs", adj = 0, cex = 0.8)
arrows(12.5,ytop1,12.5,1.2,code=2,lwd=2,angle=25,length=0.10)

text(15.5,ytop,labels="Contract \nAlteration\nIncrease \nCoverage", adj = 0, cex = 0.8)
arrows(16.2,ytop1,16.2,1.2,code=2,lwd=2,angle=25,length=0.10)

text(17.2,ytop,labels="Policy \nNonrenewal", adj = 0, cex = 0.8)
arrows(17.8,ytop1,17.8,1.2,code=2,lwd=2,angle=25,length=0.10)
text(0.8,0.6,expression(t[1]), adj = 0, cex = 0.8)
text(4.2,0.6,expression(t[2]), adj = 0, cex = 0.8)
text(9.1,0.6,expression(t[3]), adj = 0, cex = 0.8)
text(12.5,0.6,expression(t[4]), adj = 0, cex = 0.8)
text(16.0,0.6,expression(t[5]), adj = 0, cex = 0.8)
text(17.8,0.6,expression(t[6]), adj = 0, cex = 0.8)

#  Valuation Dates
text(2.7,1.8,labels="Valuation\nDate 1", adj = 0, cex = 0.8)
text(3,1,labels="X", adj = 0, cex = 0.8)
text(6.7,1.8,labels="Valuation\nDate 2", adj = 0, cex = 0.8)
text(7,1,labels="X", adj = 0, cex = 0.8)
text(10.7,1.8,labels="Valuation\nDate 3", adj = 0, cex = 0.8)
text(11,1,labels="X", adj = 0, cex = 0.8)
text(14.7,1.8,labels="Valuation\nDate 4", adj = 0, cex = 0.8)
text(15,1,labels="X", adj = 0, cex = 0.8)

#  Premium Renewals
text(4.9,1.8,labels="Policy \nRenewal\nPremium \nPaid", adj = 0, cex = 0.8)
text(5,1,labels="X", adj = 0, cex = 0.8)
text(12.9,1.8,labels="Policy \nRenewal\nPremium \nPaid", adj = 0, cex = 0.8)
text(13,1,labels="X", adj = 0, cex = 0.8)
```


## Insurance Company Operations {#S:PredModApps}

***
In this section, you learn how to:

- Describe five major operational areas of insurance companies.
- Identify the role of data and analytics opportunities within each operational area.

***

Armed with insurance data, the end goal is to use data to make decisions. We will learn more about methods of analyzing and extrapolating data in future chapters. To begin, let us think about why we want to do the analysis.
Let us take the insurer's viewpoint (not a person)
and introduce ways of bringing money in, paying it out, managing costs,
and making sure that we have enough money to meet obligations.

Specifically, in many insurance companies, it is customary to aggregate
detailed insurance processes into larger operational units; many
companies use these functional areas to segregate employee activities
and areas of responsibilities. Actuaries and other financial analysts
work within these units and use data for the following activities:

1.  **Initiating Insurance**. At this stage, the company makes a
    decision as to whether or not to take on a risk (the
    <a href="#" class="tooltip" style="color:green">*underwriting*<span style="font-size:8pt">Underwriting is the process where the company makes a decision as to
    whether or not to take on a risk.</span></a> stage) and assign an appropriate 
    premium (or rate).Insurance analytics has its actuarial roots in *ratemaking*, where
    analysts seek to determine the right price for the right risk.

2.  **Renewing Insurance**. Many contracts, particularly in general
    insurance, have relatively short durations such as 6 months or
    a year. Although there is an implicit expectation that such
    contracts will be renewed, the insurer has the opportunity to
    decline coverage and to adjust the premium. Analytics is also used
    at this policy renewal stage where the goal is to retain
    profitable customers.

3.  **Claims Management**. Analytics has long been used in (1) detecting
    and preventing claims fraud, (2) managing claim costs, including
    identifying the appropriate support for claims handling expenses, as
    well as (3) understanding excess layers for reinsurance
    and retention.

4.  **Loss Reserving**. Analytic tools are used to provide management
    with an appropriate estimate of future obligations and to quantify
    the uncertainty of those estimates.

5.  **Solvency and Capital Allocation**. Deciding on the requisite
    amount of capital and on ways of allocating capital among alternative
    investments are also important
    analytics activities. Companies must understand how much capital is
    needed so that they will have sufficient flow of cash available to
    meet their obligations. This is an important question that concerns
    not only company managers but also customers, company shareholders,
    regulatory authorities, as well as the public at large. Related to
    issues of how much capital is the question of how to allocate
    capital to differing financial projects, typically to maximize an
    investor's return. Although this question can arise at several
    levels, insurance companies are typically concerned with how to
    allocate capital to different lines of business within a firm and to
    different subsidiaries of a parent firm.

Although data represent a critical component of solvency and capital
allocation, other components including an economic framework and
the financial investments environment are also important. Because of the
background needed to address these components, we will not address
solvency and capital allocation issues further in this text.

Nonetheless, for all operating functions, we emphasize that analytics in
the insurance industry is not an exercise that a small group of analysts
can do by themselves. It requires an insurer to make significant
investments in their information technology, marketing, underwriting,
and actuarial functions. As these areas represent the primary end goals
of the analysis of data, additional background on each operational unit
is provided in the following subsections.

### Initiating Insurance

Setting the price of an insurance product can be a perplexing problem. This is in contrast to other industries such as manufacturing where the cost of a product is (relatively) known and provides a benchmark for assessing a market demand price. Similarly, in other areas of financial services, market prices are available and provide the basis
for a market-consistent pricing structure of products. However, for many lines of insurance, the cost of a product is uncertain and market prices are unavailable. Expectations of the random cost is a reasonable
place to start for a price. (If you have studies finance, then you will recall that an expectation is the optimal price for a risk-neutral insurer.) It has been traditional in insurance pricing to begin with the expected cost. Insurers then add to this  margins to account for the product's riskiness,
expenses incurred in servicing the product, and an allowance for profit/surplus of the company.

Use of expected costs as a foundation for pricing is prevalent in some lines of the insurance business. These include automobile and homeowners insurance. For these lines, analytics has served to sharpen the market by making the
calculation of the product's expected cost more precise. The increasing availability of the internet to consumers has also promoted transparency in pricing; in today's marketplace, consumers have ready access to competing quotes from a host of insurers. Insurers seek to increase their market share by refining their 
<a href="#" class="tooltip" style="color:green">risk classification<span style="font-size:8pt">Risk classification is the process of grouping policyholders into categories, or classes, where each insured in the class has a risk profile that is similar to others in the class.</span></a> systems and employing skimming the cream
underwriting strategies. Recent surveys (e.g., @survey2013) indicate
that pricing is the most common use of analytics among insurers.

*Underwriting*, the process of classifying risks into homogeneous
categories and assigning policyholders to these categories, lies at the
core of ratemaking. Policyholders within a class have similar risk
profiles and so are charged the same insurance price. This is the
concept of an actuarially fair premium; it is fair to charge
different rates to policyholders only if they can be separated by
identifiable risk factors. An early article, *Two Studies in Automobile Insurance Ratemaking* [@bailey1960], provided a catalyst to the acceptance of analytic methods in the insurance
industry. This paper addresses the problem of classification ratemaking.
It describes an example of automobile insurance that has five use
classes cross-classified with four merit rating classes. At that time,
the contribution to premiums for use and merit rating classes were
determined independently of each other. Thinking about the interacting
effects of different classification variables is a more difficult
problem.

### Renewing Insurance

Insurance is a type of financial service and, like many service
contracts, insurance coverage is often agreed upon for a limited time
period, such as six months or a year, at which time commitments are
complete. Particularly for general insurance, the need for coverage
continues and so efforts are made to issue a new contract providing
similar coverage. Renewal issues can also arise in life insurance, e.g.,
term (temporary) life insurance, although other contracts, such as life
annuities, terminate upon the insured's death and so issues of
renewability are irrelevant.

In absence of legal restrictions, at renewal the insurer has the
opportunity to:

-   accept or decline to underwrite the risk and

-   determine a new premium, possibly in conjunction with a new
    classification of the risk.

Risk classification and rating at renewal is based on two types of
information. First, as at the initial stage, the insurer has available
many rating variables upon which decisions can be made. Many variables
will not change, e.g., sex, whereas others are likely to have changed,
e.g., age, and still others may or may not change, e.g., credit score.
Second, unlike the initial stage, at renewal the insurer has available a
history of policyholder's loss experience, and this history can provide
insights into the policyholder that are not available from rating
variables. Modifying premiums with claims history is known as
*experience rating*, also sometimes referred to as *merit rating*.

Experience rating methods are either applied retrospectively or
prospectively. With retrospective methods, a refund of a portion of the
premium is provided to the policyholder in the event of favorable (to
the insurer) experience. 
<a href="#" class="tooltip" style="color:green">Retrospective premiums<span style="font-size:8pt">The process of determining the cost of an insurance policy based on the actual loss experience determined as an adjustment to the initial premium payment.</span></a> are common in life insurance arrangements (where policyholders earn <a href="#" class="tooltip" style="color:green">dividends<span style="font-size:8pt">A dividend is the refund of a portion of the premium paid by the insured from insurer surplus.</span></a> in the U.S. and bonuses in the U.K.). In general insurance, prospective methods are more common, where favorable insured experience is rewarded
through a lower renewal premium.

Claims history can provide information about a policyholder's risk
appetite. For example, in personal lines it is common to use a variable
to indicate whether or not a claim has occurred in the last three years.
As another example, in a commercial line such as worker's compensation,
one may look to a policyholder's average claim over the last three
years. Claims history can reveal information that is hidden (to the
insurer) about the policyholder.

### Claims and Product Management

In some of areas of insurance, the process of paying claims for insured
events is relatively straightforward. For example, in life insurance, a
simple death certificate is all that is needed as the benefit amount is
provided in the contract terms. However, in non-life areas such as
property and casualty insurance, the process is much more complex. Think
about even a relatively simple insured event such as automobile
accident. Here, it is often helpful to determine which party is at
fault, one needs to assess damage to all of the vehicles and people
involved in the incident, both insured and non-insured, the expenses
incurred in assessing the damages, and so forth. The process of
determining coverage, legal liability, and settling claims is known as <a href="#" class="tooltip" style="color:green">*claims adjustment*.<span style="font-size:8pt">Claims adjustment is the process of determining coverage, legal liability, and settling claims.</span></a>


Insurance managers sometimes use the phrase 
<a href="#" class="tooltip" style="color:green">*claims leakage*<span style="font-size:8pt">Claims leakage represents money lost through claims management inefficiencies.</span></a> to mean
dollars lost through claims management inefficiencies. There are many
ways in which analytics can help manage the claims process,
@SASsurvey. Historically, the most important has been fraud detection.
The claim adjusting process involves reducing information asymmetry (the
claimant knows exactly what happened; the company knows some of what
happened). Mitigating fraud is an important part of claims management
process.

Fraud detection is only one aspect of managing claims. More broadly, one can think about claims management as consisting of the following components:

-   **Claims triaging**. Just as in the medical world, early
    identification and appropriate handling of high cost claims
    (patients, in the medical world), can lead to dramatic
    company savings. For example, in workers compensation, insurers look
    to achieve early identification of those claims that run the risk of
    high medical costs and a long payout period. Early intervention into
    those cases could give insurers more control over the handling of
    the claim, the medical treatment, and the overall costs with an
    earlier return-to-work.

-   **Claims processing**. The goal is to use analytics to identify routine situations that are anticipated to have
    small payouts and more complex situations. More complex situations may require more experienced adjusters and
    legal assistance to appropriately handle claims with high potential payouts.
    
-   **Adjustment decisions**. Once a complex claim has been identified and assigned to an 
    <a href="#" class="tooltip" style="color:green">adjuster<span style="font-size:8pt">An adjuster is a person who investigates claims and recommends settlement
    options based on estimates of damage and insurance policies held.</span></a>,   
    analytic driven routines can be established to aid subsequent decision-making   
    processes. Such processes can also be helpful for adjusters in developing case
    reserves, an estimate of the insurer's future liability. This is an important 
    input to the insurer's loss reserves, described in Section \@ref(S:Reserving).

In addition to the insured's reimbursement for insured losses, the insurer also needs to be concerned with another source of revenue outflow, expenses. <a href="#" class="tooltip" style="color:green">Loss adjustment expenses<span style="font-size:8pt">Loss adjustment expenses are costs to the insurer that are directly attributable to settling a claims. For example, the cost of an adjuster is someone who assess the claim cost or a lawyer who becomes involve in settling an insurer's legal obligation on a claim</span></a> are part of an insurer's cost of managing claims. Analytics can be used to reduce expenses
directly related to claims handling (<a href="#" class="tooltip" style="color:green">allocated<span style="font-size:8pt">Allocated loss adjustment expenses, sometimes known by the acronym ALEA, are costs that can be directly attributed to settling a claim; for example, the cost of an adjuster</span></a>) as well as general staff time for overseeing the claims processes (<a href="#" class="tooltip" style="color:green">unallocated<span style="font-size:8pt">Unallocated loss adjustment expenses are costs that can only be indirectly attributed to claim settlement; for example, the cost of an office to support claims staff</span></a>). The insurance industry has high operating costs relative to other portions of the financial services sectors.

In addition to claims payments, there are many other ways in which
insurers use to data to manage their products. We have already discussed
the need for analytics in underwriting, that is, risk classification at
the initial acquisition stage. Insurers are also interested in which
policyholders elect to renew their contract and, as with other products,
monitor customer loyalty.

Analytics can also be used to manage the portfolio, or collection, of
risks that an insurer has acquired. When the risk is initially obtained,
the insurer's risk can be managed by imposing contract parameters that
modify contract payouts. Chapters \@ref(C:Severity) and \@ref(C:PortMgt) describe common modifications including <a href="#" class="tooltip" style="color:green">coinsurance,<span style="font-size:8pt">Coinsurance is an arrangement whereby the insured and insurer share the covered losses. Typically, a coinsurance parameter specified means that both parties receive a proportional share, e.g., 50%, of the loss.</span></a>
<a href="#" class="tooltip" style="color:green">deductibles,<span style="font-size:8pt">A deductible is a parameter specified in the contract. Typically, losses below the deductible are paid by the policyholder whereas losses in excess of the deductible are the insurer's responsibility (subject to policy limits and coinsurance).</span></a> and 
<a href="#" class="tooltip" style="color:green">policy upper limits.<span style="font-size:8pt">A policy limit is the maximum value covered by a policy.</span></a>

 
After the contract has been agreed upon with an insured, the insurer may
still modify its net obligation by entering into a reinsurance
agreement. This type of agreement is with a reinsurer, an insurer of an
insurer. It is common for insurance companies to purchase insurance on
its portfolio of risks to gain protection from unusual events, just as
people and other companies do.

### Loss Reserving {#S:Reserving}

An important feature that distinguishes insurance from other sectors of
the economy is the timing of the exchange of considerations. In
manufacturing, payments for goods are typically made at the time of a
transaction. In contrast, for insurance, money received from a customer
occurs in advance of benefits or services; these are rendered at a later
date. This leads to the need to hold a reservoir of wealth to meet
future obligations in respect to obligations made. The size of this
reservoir of wealth, and the importance of ensuring its adequacy, is a major concern for the
insurance industry.

Setting aside money for unpaid claims is known as <a href="#" class="tooltip" style="color:green">*loss reserving*;<span style="font-size:8pt">A loss reserve is an estimate of liability indicating the amount the insurer expects to pay for claims that have not yet been realized. This includes losses incurred but not yet reported (IBNR) and those claims that have been reported claims that haven't been paid (known by the acronym RBNS for reported but not settled).</span></a> in some jurisdictions, reserves are also known as *technical provisions*. We saw in Figure \@ref(fig:StochOperations) several times at which a company summarizes its financial position; these times are known as 
<a href="#" class="tooltip" style="color:green">*valuation dates.*<span style="font-size:8pt">A valuation date is the date at which a company summarizes its financial position, typically quarterly or annually.</span></a> Claims that arise prior to valuation dates have typically been or are about to paid; claims in the future of these valuation dates are unknown. A company must estimate these outstanding liabilities when determining its financial strength. Accurately determining loss reserves is important to insurers for many reasons.

1.  Loss reserves represent an anticipated claim that the insurer owes its customers.
    Under-reserving may result in a failure to meet claim liabilities.
    Conversely, an insurer with excessive reserves may present a weaker
    financial position than it truly has.

2.  Reserves provide an estimate for the unpaid cost of insurance that
    can be used for pricing contracts.

3.  Loss reserving is required by laws and regulations. The public has a
    strong interest in the financial strength of insurers.

4.  In addition to insurance company management and regulators,
    other stakeholders such as investors and customers make decisions
    that depend on company loss reserves.

Loss reserving is a topic where there are substantive differences
between life and general (also known as property and casualty, or
non-life), insurance. In life insurance, the severity (amount of loss)
is often not a source of uncertainty as payouts are specified in the
contract. The frequency, driven by mortality of the insured, is a
concern. However, because of the length of time for settlement of life
insurance contracts, the time value of money uncertainty as measured
from issue to date of death can dominate frequency concerns. For
example, for an insured who purchases a life contract at age 25, it
would not be unusual for the contract to still be open in 75 years time.
See, for example, @bowers1986actuarial or @dickson2013actuarial for
introductions to reserving for life insurance.

## Case Study: Wisconsin Property Fund {#S:LGPIF}

***
In this section, we use the Wisconsin Property Fund as a case study. You learn how to:

- Describe how data generating events can produce data of interest to insurance analysts.
- Identify each type of variable.
- Produce relevant summary statistics for each variable.
- Describe how these summary statistics can be used in each of the major operational areas of an insurance company.

***




Let us illustrate the kind of data under consideration and the goals
that we wish to achieve by examining the Local Government Property
Insurance Fund (LGPIF), an insurance pool administered by the Wisconsin
Office of the Insurance Commissioner. The LGPIF was established to
provide property insurance for local government entities that include
counties, cities, towns, villages, school districts, and library boards.
The fund insures local government property such as government buildings,
schools, libraries, and motor vehicles. The fund covers all property
losses except those resulting from flood, earthquake, wear and tear,
extremes in temperature, mold, war, nuclear reactions, and embezzlement
or theft by an employee.

The property fund covers over a thousand local government entities who
pay approximately \$25 million in premiums each year and receive
insurance coverage of about \$75 billion. State government buildings are
not covered; the LGPIF is for local government entities that have
separate budgetary responsibilities and who need insurance to moderate
the budget effects of uncertain insurable events. Coverage for local
government property has been made available by the State of Wisconsin
since 1911.

### Fund Claims Variables: Frequency and Severity {#S:OutComes}

At a fundamental level, insurance companies accept premiums in exchange for promises to compensate a policyholder upon the occurrence
of an insured event. <a href="#" class="tooltip" style="color:green">*Indemnification*<span style="font-size:8pt">Indemnification is the compensation provided by the insurer.</span></a> is the compensation provided by the insurer for incurred hurt, loss, or damage that is covered by the policy. This compensation is also known as a *claim*. The extent of the payout, known as the *severity*, is a key financial expenditure for an insurer.

In terms of money outgo, an insurer is indifferent to having ten claims of 100 when compared to one claim of 1,000. Nonetheless, it is common for
insurers to study how often claims arise, known as the *frequency* of
claims. The frequency is important for expenses, but it also influences
contractual parameters (such as deductibles and policy limits that are described later) that are
written on a per occurrence basis, is routinely monitored by insurance
regulators, and can be a key driver in the overall indemnification
obligation of the insurer. We shall consider the frequency and severity as the two main outcome variables that we
wish to understand, model, and manage.

To illustrate, in 2010 there were 1,110 policyholders in the property fund who experienced a total of 1,377 claims. Table \@ref(tab:Frequency2010) shows the distribution. Almost two-thirds (0.637) of the policyholders did not have any claims and an additional 18.8% had only one claim. The remaining 17.5%
(=1 - 0.637 - 0.188) had more than one claim; the policyholder with the highest number recorded 239 claims. The average number of claims for
this sample was 1.24 (=1377/1110).

  Type                                                                                             
  ------------- ------ ------- ------- ------- ------- ------- ------- ------- ------- ----------- -------
  Number             0       1       2       3       4       5       6       7       8   9 or more     Sum
  Count            707     209      86      40      18      12       9       4       6          19   1,110
  Claims             0     209     172     120      72      60      54      28      48         617   1,377
  Proportion     0.637   0.188   0.077   0.036   0.016   0.011   0.008   0.004   0.005       0.017   1.000
  ------------- ------ ------- ------- ------- ------- ------- ------- ------- ------- ----------- -------

Table: (\#tab:Frequency2010)  2010 Claims Frequency Distribution
  

<h5 style="text-align: center;"><a id="display.T:Frequency.1" href="javascript:togglecode
('display.T:Frequency.2','display.T:Frequency.1');"><i><strong>R Code for Frequency Table</strong></i></a> </h5>
<div id="display.T:Frequency.2" style="display: none">

``` 
Insample <- read.csv("Insample.csv", header=T,  na.strings=c("."), stringsAsFactors=FALSE)
Insample2010 <- subset(Insample, Year==2010)
table(Insample2010$Freq)
```
</div>  
  

For the severity distribution, one common approach is to examine the
distribution of the sample of 1,377 claims. However, another common
approach is to examine the distribution of the average claims of those
policyholders with claims. In our 2010 sample, there were 403
(=1110-707) such policyholders. For 209 of these policyholders with one
claim, the average claim equals the only claim they experienced. For the
policyholder with highest frequency, the average claim is an average
over 239 separately reported claim events. This average is also known as the 
<a href="#" class="tooltip" style="color:green">*pure premium*<span style="font-size:8pt">Pure premium is the total severity divided by the number of claims. It does not include insurance company expenses, premium taxes, contingencies, nor an allowance for profits. Also called loss costs. Some definitions include allocated loss adjustment expenses (ALAE).</span></a>
 or *loss cost*.

Table \@ref(tab:Severity2010) summarizes the sample distribution of average
severities from the 403 policyholders; it shows that the average claim
amount was 56,330 (all amounts are in US Dollars). However, the average
gives only a limited look at the distribution. More information can be
gleaned from the summary statistics which show a very large claim in the
amount of 12,920,000. Figure \@ref(fig:SeverityFig) provides further
information about the distribution of sample claims, showing a
distribution that is dominated by this single large claim so that the
histogram is not very helpful. Even when removing the large claim, you
will find a distribution that is skewed to the right. A generally
accepted technique is to work with claims in logarithmic units
especially for graphical purposes; the corresponding figure in the
right-hand panel is much easier to interpret.

  --------- ---------- -------- -------- ---------- ------------
                 First                        Third 
    Minimum   Quartile   Median     Mean   Quartile      Maximum
  --------- ---------- -------- -------- ---------- ------------
        167      2,226    4,951   56,330     11,900   12,920,000
  --------- ---------- -------- -------- ---------- ------------

Table: (\#tab:Severity2010)  2010 Average Severity Distribution
  
 


```{r SeverityFig, echo=FALSE, fig.cap='Distribution of Positive Average Severities', out.width='80%', fig.asp=.75, fig.align='center'}
Insample <- read.csv("Data/PropertyFundInsample.csv", header=T, na.strings=c("."), stringsAsFactors=FALSE)
Insample2010 <- subset(Insample, Year==2010)
InsamplePos2010 <- subset(Insample2010, yAvg>0)
par(mfrow=c(1, 2))
hist(InsamplePos2010$yAvg, main="", xlab="Average Claims")
hist(log(InsamplePos2010$yAvg), main="", xlab="Logarithmic Average Claims")
```

<h5 style="text-align: center;"><a id="display.SeverityFig.1" href="javascript:togglecode('display.SeverityFig.2','display.SeverityFig.1');"><i><strong>R Code for Severity Distribution Table and Figures</strong></i></a> </h5>
<div id="display.SeverityFig.2" style="display: none">

``` 
Insample <- read.csv("Data/PropertyFundInsample.csv", header=T, na.strings=c("."), stringsAsFactors=FALSE)
Insample2010 <- subset(Insample, Year==2010)
InsamplePos2010 <- subset(Insample2010, yAvg>0)
# Table
summary(InsamplePos2010$yAvg)
length(InsamplePos2010$yAvg)
# Figures
par(mfrow=c(1, 2))
hist(InsamplePos2010$yAvg, main="", xlab="Average Claims")
hist(log(InsamplePos2010$yAvg), main="", xlab="Logarithmic Average Claims")
```
</div>


### Fund Rating Variables {#S:FundVariables}

Developing models to represent and manage the two outcome variables,
frequency and severity, is the focus of the early chapters of this text.
However, when actuaries and other financial analysts use those models,
they do so in the context of external variables. In general
statistical terminology, one might call these explanatory or predictor
variables; there are many other names in statistics, economics,
psychology, and other disciplines. Because of our insurance focus, we
call them *rating variables* as they will be useful in setting insurance
rates and premiums.

We earlier considered a sample of 1,110 observations which may seem like
a lot. However, as we will see in our forthcoming applications, because
of the preponderance of zeros and the skewed nature of claims, actuaries
typically yearn for more data. One common approach that we adopt here is
to examine outcomes from multiple years, thus increasing the sample
size. We will discuss the strengths and limitations of this strategy
later but, at this juncture, just want to show the reader how it works.

Specifically, Table \@ref(tab:CoverageBCIM) shows that we now consider
policies over five years of data, years 2006, ..., 2010, inclusive. The
data begins in 2006 because there was a shift in claim coding in 2005 so
that comparisons with earlier years are not helpful. To mitigate the
effect of open claims, we consider policy years prior to 2011. An open
claim means that all of the obligations are not known at the time of the
analysis; for some claims, such an injury to a person in an auto
accident or in the workplace, it can take years before costs are fully
known.

Table \@ref(tab:CoverageBCIM) shows that the average claim varies over time,
especially with the high 2010 value due to a single large claim [^1.1]. The
total number of policyholders is steadily declining and, conversely, the
coverage is steadily increasing. The coverage variable is the amount of
coverage of the property and contents. Roughly, you can think of it as
the maximum possible payout of the insurer. For our immediate purposes,
it is our first rating variable. Other things being equal, we would
expect that policyholders with larger coverage will have larger claims.
We will make this vague idea much more precise as we proceed.


  ------------------------------------------------------------------
  Year            Average    Average         Average       Number of
                Frequency   Severity        Coverage   Policyholders
  -----------   ---------   --------      ----------   -------------
  2006              0.951      9,695      32,498,186           1,154
  
  2007              1.167      6,544      35,275,949           1,138
  
  2008              0.974      5,311      37,267,485           1,125
  
  2009              1.219      4,572      40,355,382           1,112
  
  2010              1.241     20,452      41,242,070           1,110
  ------------------------------------------------------------------

  Table: (\#tab:CoverageBCIM) Building and Contents Claims Summary

[^1.1]: Note that the average severity in Table \@ref(tab:CoverageBCIM) differs from that reported in Table \@ref(tab:Severity2010). This is because the former includes policyholders with zero claims where as the latter does not. This is an important distinction that we will address in later portions of the text. 

<h5 style="text-align: center;"><a id="display.CoverageBC.1" href="javascript:togglecode('display.CoverageBC.2','display.CoverageBC.1');"><i><strong>R Code for Building and Contents Claims Summary</strong></i></a> </h5>
<div id="display.CoverageBC.2" style="display: none">

``` 
Insample <- read.csv("Data/PropertyFundInsample.csv", header=T, na.strings=c("."), stringsAsFactors=FALSE)
library(doBy)
T1A <- summaryBy(Freq ~ Year, data = Insample, 
   FUN = function(x) { c(m = mean(x), num=length(x)) } )
T1B <- summaryBy(yAvg    ~ Year, data = Insample,   
   FUN = function(x) { c(m = mean(x), num=length(x)) } )
T1C <- summaryBy(BCcov    ~ Year, data = Insample,   
   FUN = function(x) { c(m = mean(x), num=length(x)) } )
Table1In <- cbind(T1A[1],T1A[2],T1B[2],T1C[2],T1A[3])
names(Table1In) <- c("Year", "Average Frequency","Average Severity", "Average","Number of Policyholders")
Table1In
```
</div>


For a different look at this data, Table \@ref(tab:DeductCov)
summarizes the distribution of our two outcomes, frequency and claims
amount. In each case, the average exceeds the median, suggesting that
the two distributions are right-skewed. In addition, the table
summarizes our continuous rating variables, coverage and deductible
amount. The table also suggests that these variables also have
right-skewed distributions.

  ------------------ --------- -------- --------- ------------
                       Minimum   Median   Average      Maximum    
  ------------------ --------- -------- --------- ------------
  Claim Frequency            0        0     1.109          263
  
  Claim Severity             0        0     9,292   12,922,218
  
  Deductible               500    1,000     3,365      100,000 
  
  Coverage (000's)       8.937   11,354    37,281    2,444,797    
  ------------------ --------- -------- --------- ------------
  
Table: (\#tab:DeductCov) Summary of Claim Frequency and Severity, Deductibles, and Coverages
  

<h5 style="text-align: center;"><a id="display.DeductCov.1" href="javascript:togglecode('display.DeductCov.2','display.DeductCov.1');"><i><strong>R Code for Summary of Claim Frequency and Severity, Deductibles, and Coverages</strong></i></a> </h5>
<div id="display.DeductCov.2" style="display: none">

``` 
Insample <- read.csv("Data/PropertyFundInsample.csv", header=T, na.strings=c("."), stringsAsFactors=FALSE)
t1<- summaryBy(Insample$Freq ~ 1, data = Insample, 
   FUN = function(x) { c(ma=min(x), m1=median(x),m=mean(x),mb=max(x)) } )
names(t1) <- c("Minimum", "Median","Average", "Maximum")
t2 <- summaryBy(Insample$yAvg ~ 1, data = Insample, 
   FUN = function(x) { c(ma=min(x), m1=median(x), m=mean(x),mb=max(x)) } )
names(t2) <- c("Minimum", "Median","Average", "Maximum")
t3 <- summaryBy(Deduct ~ 1, data = Insample, 
   FUN = function(x) { c(ma=min(x), m1=median(x), m=mean(x),mb=max(x)) } )
names(t3) <- c("Minimum", "Median","Average", "Maximum")
t4 <- summaryBy(BCcov/1000 ~ 1, data = Insample, 
   FUN = function(x) { c(ma=min(x), m1=median(x), m=mean(x),mb=max(x)) } )
names(t4) <- c("Minimum", "Median","Average", "Maximum")
Table2 <- rbind(t1,t2,t3,t4)
Table2a <- round(Table2,3)
Rowlable <- rbind("Claim Frequency","Claim Severity","Deductible","Coverage (000's)")
Table2aa <- cbind(Rowlable,as.matrix(Table2a))
Table2aa
```
</div>  
  

The following display describes the rating variables considered in this
chapter. To handle the skewness, we henceforth focus on logarithmic
transformations of coverage and deductibles. To get a sense of the
relationship between the non-continuous rating variables and claims,
Table \@ref(tab:ClaimRateVar) relates the claims outcomes to these
categorical variables. Table \@ref(tab:ClaimRateVar) suggests substantial
variation in the claim frequency and average severity of the claims by
entity type. It also demonstrates higher frequency and severity for the
 ${\tt Fire5}$ variable and the reverse for the ${\tt NoClaimCredit}$ variable.
The relationship for the ${\tt Fire5}$ variable is counter-intuitive in that
one would expect lower claim amounts for those policyholders in areas
with better public protection (when the protection code is five or
less). Naturally, there are other variables that influence this
relationship. We will see that these background variables are accounted
for in the subsequent multivariate regression analysis, which yields an
intuitive, appealing (negative) sign for the ${\tt Fire5}$ variable.

Description of Rating Variables 
$${\small \begin{matrix}
\begin{array}{ l | l}
\hline
Variable 	& Description \\
\hline
\text{EntityType} 	& \text{Categorical variable that is one of six types:  (Village, City,} \\
& ~~~~ \text{County, Misc, School, or Town)} \\
\text{LnCoverage}	& \text{Total building and content coverage, in logarithmic millions of dollars}\\
\text{LnDeduct}   	& \text{Deductible, in logarithmic dollars} \\
\text{AlarmCredit}  & \text{Categorical variable that is one of four types:  (0, 5, 10, or 15)} \\
 &  ~~~~   \text{for automatic smoke alarms in main rooms} \\
\text{NoClaimCredit}  	& \text{Binary variable to indicate no claims in the past two years} \\
\text{Fire5 }			& \text{Binary variable to indicate the fire class is below 5} \\
& ~~~~ \text{(The range of fire class is 0 to 10)} \\
\hline
\end{array}
\end{matrix}}$$
<!-- <Description of Rating Variables>(\#tab:VarDesc)  -->



  --------------------- ----------- ----------- ----------
                          Number of       Claim    Average
  Variable                 Policies   Frequency   Severity
  --------------------- ----------- ----------- ----------
  *EntityType* 
  
  Village                     1,341       0.452     10,645
  
  City                          793       1.941     16,924
  
  County                        328       4.899     15,453
  
  Misc                          609       0.186     43,036
  
  School                      1,597       1.434     64,346
  
  Town                          971       0.103     19,831
  
  *Fire*
  
  Fire5=0                     2,508       0.502     13,935
  
  Fire5=1                     3,131       1.596     41,421
  
  *No Claims Credit*
  
  NoClaimCredit=0             3,786       1.501     31,365
  
  NoClaimCredit=1             1,853       0.310     30,499
  
  **Total**                   5,639       1.109     31,206
  --------------------- ----------- ----------- ----------

 Table: (\#tab:ClaimRateVar) Claims Summary by Entity Type, Fire Class, and No Claim Credit
  
  
<h5 style="text-align: center;"><a id="display.ClaimRateVar.1" href="javascript:togglecode('display.ClaimRateVar.2','display.ClaimRateVar.1');"><i><strong>R Code for Claims Summary by Entity Type, Fire Class, and No Claim Credit</strong></i></a> </h5>
<div id="display.ClaimRateVar.2" style="display: none">

``` 
ByVarSumm<-function(datasub){
  tempA <- summaryBy(Freq    ~ 1 , data = datasub,   
     FUN = function(x) { c(m = mean(x), num=length(x)) } )
  datasub1 <-  subset(datasub, yAvg>0)
  tempB <- summaryBy(yAvg   ~ 1, data = datasub1,FUN = function(x) { c(m = mean(x)) } )
  tempC <- merge(tempA,tempB,all.x=T)[c(2,1,3)]
  tempC1 <- as.matrix(tempC)
  return(tempC1)
  }
datasub <-  subset(Insample, TypeVillage == 1);   
t1 <- ByVarSumm(datasub)
datasub <-  subset(Insample, TypeCity == 1);      
t2 <- ByVarSumm(datasub)
datasub <-  subset(Insample, TypeCounty == 1);   
t3 <- ByVarSumm(datasub)
datasub <-  subset(Insample, TypeMisc == 1);      
t4 <- ByVarSumm(datasub)
datasub <-  subset(Insample, TypeSchool == 1);    
t5 <- ByVarSumm(datasub)
datasub <-  subset(Insample, TypeTown == 1);      
t6 <- ByVarSumm(datasub)
datasub <-  subset(Insample, Fire5 == 0);                      
t7 <- ByVarSumm(datasub)
datasub <-  subset(Insample, Fire5 == 1);                      
t8 <- ByVarSumm(datasub)
datasub <-  subset(Insample, Insample$NoClaimCredit == 0);
t9 <- ByVarSumm(datasub)
datasub <-  subset(Insample, Insample$NoClaimCredit == 1);
t10 <- ByVarSumm(datasub)
t11 <- ByVarSumm(Insample)

Tablea <- rbind(t1,t2,t3,t4,t5,t6,t7,t8,t9,t10,t11)
Tableaa <- round(Tablea,3)
Rowlable <- rbind("Village","City","County","Misc","School",
          "Town","Fire5--No","Fire5--Yes","NoClaimCredit--No",
        "NoClaimCredit--Yes","Total")
Table4 <- cbind(Rowlable,as.matrix(Tableaa))
Table4
```
</div>  
    

Table \@ref(tab:RateAlarmCredit) shows the claims experience by alarm credit.
It underscores the difficulty of examining variables individually. For
example, when looking at the experience for all entities, we see that
policyholders with no alarm credit have on average lower frequency and
severity than policyholders with the highest (15%, with 24/7 monitoring
by a fire station or security company) alarm credit. In particular, when
we look at the entity type School, the frequency is 0.422 and the
severity 25,523 for no alarm credit, whereas for the highest alarm level
it is 2.008 and 85,140. This may simply imply that entities with more
claims are the ones that are likely to have an alarm system. Summary
tables do not examine multivariate effects; for example, Table
\@ref(tab:ClaimRateVar) ignores the effect of size (as we measure through
coverage amounts) that affect claims.

  -----------------------------------------------------------------------------
                    AC0        AC0        AC0         AC5        AC5        AC5
  Entity          Claim       Avg.       Num.       Claim       Avg.       Num.
  Type        Frequency   Severity   Policies   Frequency   Severity   Policies
  --------- ----------- ---------- ---------- ----------- ---------- ----------
  Village         0.326     11,078        829       0.278      8,086         54
  
  City            0.893      7,576        244       2.077      4,150         13
  
  County          2.140     16,013         50           -          -          1
  
  Misc            0.117     15,122        386       0.278     13,064         18
  
  School          0.422     25,523        294       0.410     14,575        122
  
  Town            0.083     25,257        808       0.194      3,937         31
  
  Total           0.318     15,118      2,611       0.431     10,762        239
  -----------------------------------------------------------------------------

  Table: (\#tab:RateAlarmCredit) Claims Summary by Entity Type and Alarm Credit (AC) Category
  
   

  -----------------------------------------------------------------------------
                   AC10      AC10        AC10        AC15       AC15       AC15
  Entity          Claim       Avg.       Num.       Claim       Avg.       Num.
  Type        Frequency   Severity   Policies   Frequency   Severity   Policies
  
  --------- ----------- ---------- ---------- ----------- ---------- ----------
  Village         0.500      8,792         50       0.725     10,544        408
  
  City            1.258      8,625         31       2.485     20,470        505
  
  County          2.125     11,688          8       5.513     15,476        269
  
  Misc            0.077      3,923         26       0.341     87,021        179
  
  School          0.488     11,597        168       2.008     85,140      1,013
  
  Town            0.091      2,338         44       0.261      9,490         88
  
  Total           0.517     10,194        327       2.093     41,458      2,462
  -----------------------------------------------------------------------------

  : Claims Summary by Entity Type and Alarm Credit (AC) Category
  



<h5 style="text-align: center;"><a id="display.RateAlarmCredit.1" href="javascript:togglecode('display.RateAlarmCredit.2','display.RateAlarmCredit.1');"><i><strong>R Code for Claims Summary by Entity Type and Alarm Credit
  Category</strong></i></a> </h5>
<div id="display.RateAlarmCredit.2" style="display: none">

``` 
#Claims Summary by Entity Type and Alarm Credit
ByVarSumm<-function(datasub){
  tempA <- summaryBy(Freq    ~ AC00 , data = datasub,   
                     FUN = function(x) { c(m = mean(x), num=length(x)) } )
  datasub1 <-  subset(datasub, yAvg>0)
  if(nrow(datasub1)==0) { n<-nrow(datasub)
    return(c(0,0,n))
  } else 
  {
    tempB <- summaryBy(yAvg   ~ AC00, data = datasub1,
                       FUN = function(x) { c(m = mean(x)) } )
    tempC <- merge(tempA,tempB,all.x=T)[c(2,4,3)]
    tempC1 <- as.matrix(tempC)
    return(tempC1)
  }
}
AlarmC <- 1*(Insample$AC00==1) + 2*(Insample$AC05==1)+ 3*(Insample$AC10==1)+ 4*(Insample$AC15==1)
ByVarCredit<-function(ACnum){
datasub <-  subset(Insample, TypeVillage == 1 & AlarmC == ACnum); 
  t1 <- ByVarSumm(datasub)
datasub <-  subset(Insample, TypeCity == 1 & AlarmC == ACnum);      
  t2 <- ByVarSumm(datasub)
datasub <-  subset(Insample, TypeCounty == 1 & AlarmC == ACnum);   
  t3 <- ByVarSumm(datasub)
datasub <-  subset(Insample, TypeMisc == 1 & AlarmC == ACnum);
  t4 <- ByVarSumm(datasub)
datasub <-  subset(Insample, TypeSchool == 1 & AlarmC == ACnum);    
  t5 <- ByVarSumm(datasub)
datasub <-  subset(Insample, TypeTown == 1 & AlarmC ==ACnum);      
  t6 <- ByVarSumm(datasub)
datasub <-  subset(Insample, AlarmC == ACnum);  
  t7 <- ByVarSumm(datasub)
Tablea <- rbind(t1,t2,t3,t4,t5,t6,t7)
Tableaa <- round(Tablea,3)
Rowlable <- rbind("Village","City","County","Misc","School",
                  "Town","Total")
Table4 <- cbind(Rowlable,as.matrix(Tableaa))
}
Table4a <- ByVarCredit(1)    #Claims Summary by Entity Type and Alarm Credit==00
Table4b <- ByVarCredit(2)    #Claims Summary by Entity Type and Alarm Credit==05 
Table4c <- ByVarCredit(3)    #Claims Summary by Entity Type and Alarm Credit==10
Table4d <- ByVarCredit(4)    #Claims Summary by Entity Type and Alarm Credit==15

```
</div>  
      
  

### Fund Operations

We have now seen the Fund's two outcome variables, a count variable for
the number of claims and a continuous variable for the claims amount. We
have also introduced a continuous rating variable, coverage, discrete
quantitative variable, (logarithmic) deductibles, two binary rating
variables, no claims credit and fire class, as well as two categorical
rating variables, entity type and alarm credit. Subsequent chapters will
explain how to analyze and model the distribution of these variables and
their relationships. Before getting into these technical details, let us
first think about where we want to go. General insurance company
functional areas are described in Section \@ref(S:PredModApps); let us now
think about how these areas might apply in the context of the property
fund.

####Initiating Insurance {-}

Because this is a government sponsored fund, we do not have to worry
about selecting good or avoiding poor risks; the fund is not allowed to
deny a coverage application from a qualified local government entity. If
we do not have to underwrite, what about how much to charge?

We might look at the most recent experience in 2010, where the total
fund claims were approximately 28.16 million USD
($=1377 \text{ claims} \times 20452 \text{ average severity}$). Dividing
that among 1,110 policyholders, that suggests a rate of 24,370 (
$\approx$ 28,160,000/1110). However, 2010 was a bad year; using the same
method, our premium would be much lower based on 2009 data. This swing
in premiums would defeat the primary purpose of the fund, to allow for a
steady charge that local property managers could utilize in their
budgets.

Having a single price for all policyholders is nice but hardly seems
fair. For example, Table \@ref(tab:ClaimRateVar) suggests that Schools have
much higher claims than other entities and so should pay more. However,
simply doing the calculation on an entity by entity basis is not right
either. For example, we saw in Table \@ref(tab:RateAlarmCredit) that had we
used this strategy, entities with a 15% alarm credit (for good behavior,
having top alarm systems) would actually wind up paying more.

So, we have the data for thinking about the appropriate rates to charge
but will need to dig deeper into the analysis. We will explore this
topic further in Chapter \@ref(C:PremCalc) on *premium calculation fundamentals*.
Selecting appropriate risks is introduced in Chapter \@ref(C:RiskClass) on *risk
classification*.

####Renewing Insurance {-}

Although property insurance is typically a one-year contract, Table
\@ref(tab:CoverageBCIM) suggests that policyholders tend to renew; this is
typical of general insurance. For renewing policyholders, in addition to
their rating variables we have their claims history and this claims
history can be a good predictor of future claims. For example, Table
\@ref(tab:ClaimRateVar) shows that policyholders without a claim in the last
two years had much lower claim frequencies than those with at least one
accident (0.310 compared to 1.501); a lower predicted frequency
typically results in a lower premium. This is why it is common for
insurers to use variables such as ${\tt NoClaimCredit}$ in their rating. We
will explore this topic further in Chapter \@ref(C:Credibility) on *experience rating*.

####Claims Management {-}

Of course, the main story line of 2010 experience was the large claim of
over 12 million USD, nearly half the claims for that year. Are there
ways that this could have been prevented or mitigated? Are their ways
for the fund to purchase protection against such large unusual events?
Another unusual feature of the 2010 experience noted earlier was the
very large frequency of claims (239) for one policyholder. Given that
there were only 1,377 claims that year, this means that a single
policyholder had 17.4 % of the claims. This also suggests
opportunities for managing claims, the subject of Chapter \@ref(C:PortMgt).

####Loss Reserving {-}

In our case study, we look only at the one year outcomes of closed
claims (the opposite of open). However, like many lines of insurance,
obligations from insured events to buildings such as fire, hail, and the
like, are not known immediately and may develop over time. Other lines
of business, including those were there are injuries to people, take
much longer to develop. Chapter \@ref(C:LossReserves) introduces this concern and *loss
reserving*, the discipline of determining how much the insurance company
should retain to meet its obligations.

## Further Resources and Contributors {#Intro-further-reading-and-resources}

####Contributor {-}

- **Edward W. (Jed) Frees**, University of Wisconsin-Madison, is the principal author of the initial version of this chapter. Email: jfrees@bus.wisc.edu for chapter comments and suggested improvements.

This book introduces loss data analytic tools that are most relevant to actuaries and other financial risk analysts. We have also introduced you to many new insurance terms; more terms can be found at the @NAICGlossary. Here are a few reference cited in the chapter.



