
```{r echo = FALSE, message=FALSE, warning=FALSE}
chapnum = 7

```

#Premium Foundations {#C:PremiumFoundations}

*Chapter Preview.* 

***

**This chapter is being written and is not yet complete nor edited. It is here to give you a flavor of what will be in the final version.**

***


## Introduction to Ratemaking {#S:IntroductionRatemaking}

***
In this section, you learn how to:

- Describe expectations as a baseline method for determining insurance premiums
- Analyze an accounting equation for relating premiums to losses, expenses and profits
- Summarize a strategy for extending pricing to include heterogeneous risks and trends over time

***


```{r child = 'Chapters/PremiumSections/RatemakingIntroduction.Rmd'}
```

## Aggregate Ratemaking Methods {#S:AggRateMaking}

***
In this section, you learn how to:

- Define a pure premium as a loss cost as well as in terms of frequency and severity
- Calculate an indicated rate use pure premiums, expenses, and profit loadings
- Define a loss ratio
- Calculate an indicated rate change using loss ratios
- Compare the pure premium and loss ratio methods for determining premiums

***

```{r child = 'Chapters/PremiumSections/AggRatemaking.Rmd'}
```


## Pricing Principles {#S:PricingPrinciples}

1. Start with expected utility and competitive marketplace.

2. Actuarial Premium Principles

**Section 10.2.1** describes variance and standard deviation principles. Also describes coherent measures. 

3. Actuarial vs Financial Pricing. 

See @bauer2013financial  chapter in *Insurance Economics.*

## Heterogeneous Risks {#S:HeterogeneousRisks}

***
In this section, you learn how to:

- Describe insurance exposures in terms of scale distributions
- Explain an exposure in terms of common types of insurance such as auto and homeowners insurance
- Describe how rating factors can be used to create homogeneous collections of risks
- Measure the impact of a rating factor through relativities

***

```{r child = 'Chapters/PremiumSections/Heterogeneous.Rmd'}
```

## Development and Trending {#S:TrendDevelopment}

***
In this section, you learn how to:

- Define and calculate different types of exposure and premium summary measures that appear in financial reports 
- Describe the development of a claim over several payments and link that to various unpaid claim measures, including those incurred but not reported (IBNR) as well as case reserves
- Compare and contrast relative strengths and weaknesses of the pure premium and loss ratio methods for ratemaking

***


```{r child = 'Chapters/PremiumSections/TrendDevelopment.Rmd'}
```

## Selecting a Premium {#S:GiniStatistic}

***
In this section, you learn how to:

- Describe skewed distributions via a Lorenz curve and Gini index
- Define a concentration curve and the corresponding Gini statistic
- Use the concentration curve and Gini statistic for premium selection base on out-of-sample validation

***


```{r child = 'Chapters/PremiumSections/PremiumSelection.Rmd'}
```



##Further Resources and Contributors


This chapter serves as a bridge between the technical introduction of this book and an introduction to pricing and ratemaking for practicing actuaries. For readers interested in learn practical aspects of pricing, we recommend introductions by the Society of Actuaries in @friedland2013fundamentals and by the Casualty Actuarial Society in @werner2016basic. For a classic risk management introduction to pricing, see 
@niehaus2003risk. See also @finger2006risk and @frees2014frequency.


- **Edward W. (Jed) Frees**, University of Wisconsin-Madison, and **Jos&eacute;  Garrido**, Concordia University are the principal authors of the initial version of this chapter. Email: jfrees@bus.wisc.edu and/or jose.garrido@concordia.ca for chapter comments and suggested improvements.
- Chapter reviewers include: Write Jed or Jos&eacute; to add you name here.

### TS 7.A. Rate Regulation {-}

```{r child = 'Chapters/PremiumSections/RateRegulation.Rmd'}
```




