
`r chapnum = 7`

#Premium Foundations {#C:PremiumFoundations}

*Chapter Preview.* Setting prices for insurance products, premiums, is an important task for actuaries and other data analysts. This chapter introduces the foundations for pricing non-life products.



## Introduction to Ratemaking {#S:IntroductionRatemaking}

***
In this section, you learn how to:

- Describe expectations as a baseline method for determining insurance premiums
- Analyze an accounting equation for relating premiums to losses, expenses and profits
- Summarize a strategy for extending pricing to include heterogeneous risks and trends over time

***


```{r child = 'Chapters/PremiumSections/RatemakingIntroduction.Rmd'}
```

## Aggregate Ratemaking Methods {#S:AggRateMaking}

***
In this section, you learn how to:

- Define a pure premium as a loss cost as well as in terms of frequency and severity
- Calculate an indicated rate using pure premiums, expenses, and profit loadings
- Define a loss ratio
- Calculate an indicated rate change using loss ratios
- Compare the pure premium and loss ratio methods for determining premiums

***

```{r child = 'Chapters/PremiumSections/AggRatemaking.Rmd'}
```


## Pricing Principles {#S:PricingPrinciples}

***
In this section, you learn how to:

- Describe common actuarial pricing principles
- Describe properties of pricing principles
- Choose a pricing principle based on a desired property

***


```{r child = 'Chapters/PremiumSections/PremPrinciples.Rmd'}
```



## Heterogeneous Risks {#S:HeterogeneousRisks}

***
In this section, you learn how to:

- Describe insurance exposures in terms of scale distributions
- Explain an exposure in terms of common types of insurance such as auto and homeowners insurance
- Describe how rating factors can be used to account for the heterogeneity among risks in a collection
- Measure the impact of a rating factor through relativities

***

```{r child = 'Chapters/PremiumSections/Heterogeneous.Rmd'}
```

## Development and Trending {#S:TrendDevelopment}

***
In this section, you learn how to:

- Define and calculate different types of exposure and premium summary measures that appear in financial reports 
- Describe the development of a claim over several payments and link that to various unpaid claim measures, including those incurred but not reported (IBNR) as well as case reserves
- Compare and contrast relative strengths and weaknesses of the pure premium and loss ratio methods for ratemaking

***


```{r child = 'Chapters/PremiumSections/TrendDevelopment.Rmd'}
```

## Selecting a Premium {#S:GiniStatistic}

***
In this section, you learn how to:

- Describe skewed distributions via a Lorenz curve and Gini index
- Define a concentration curve and the corresponding Gini statistic
- Use the concentration curve and Gini statistic for premium selection base on out-of-sample validation

***


```{r child = 'Chapters/PremiumSections/PremiumSelection.Rmd'}
```



##Further Resources and Contributors


This chapter serves as a bridge between the technical introduction of this book and an introduction to pricing and ratemaking for practicing actuaries. For readers interested in learn practical aspects of pricing, we recommend introductions by the Society of Actuaries in @friedland2013fundamentals and by the Casualty Actuarial Society in @werner2016basic. For a classic risk management introduction to pricing, see @niehaus2003risk. See also @finger2006risk and @frees2014frequency.

@buhlmann1985premium was the first in the academic literature to argue that pricing should be done first at the portfolio level (he referred to this as a *top down* approach) which would be subsequently reconciled with pricing of individual contracts. See also the discussion in @kaas2008modern, Chapter 5.

For more background on pricing principles, a classic treatment is by @gerber1979introduction with a more modern approach in 
@kaas2008modern. For more discussion of a pricing from a financial economics viewpoint, see @bauer2013financial.


- **Edward W. (Jed) Frees**, University of Wisconsin-Madison, and **Jos&eacute;  Garrido**, Concordia University are the principal authors of the initial version of this chapter. Email: jfrees@bus.wisc.edu and/or jose.garrido@concordia.ca for chapter comments and suggested improvements.
- Chapter reviewers include: Write Jed or Jos&eacute; to add you name here.

### TS 7.A. Rate Regulation {-}

```{r child = 'Chapters/PremiumSections/RateRegulation.Rmd'}
```




