#Simulation {#C:Simulation}

Simulation is a computer-based, computationally intensive, method of
solving difficult problems, such as analyzing business processes.
Instead of creating physical processes and experimenting with them in
order to understand their operational characteristics, a simulation
study is based on a computer representation - it considers various
hypothetical conditions as inputs and summarizes the results. Through
simulation, a vast number of hypothetical conditions can be quickly and
inexpensively examined. Performing the same analysis with a physical
system is not only expensive and time-consuming but, in many cases,
impossible. A drawback of simulation is that computer models are not
perfect representations of business processes.

There are three basic steps for producing a simulation study:

-   Generating approximately independent realizations that are uniformly
    distributed

-   Transforming the uniformly distributed realizations to observations
    from a probability distribution of interest

-   With the generated observations as inputs, designing a structure to
    produce interesting and reliable results.

Designing the structure can be a difficult step, where the degree of
difficulty depends on the problem being studied. There are many
resources, including this tutorial, to help the actuary with the first
two steps.

## Generating Independent Uniform Observations

We begin with a historically prominent method.

**Linear Congruential Generator.** To generate a sequence of random
numbers, start with $B_0$, a starting value that is known as a “seed.”
Update it using the recursive relationship
$$B_{n+1} = a B_n + c  \text{ modulo }m, ~~ n=0, 1, 2, \ldots .$$
This algorithm is called a *linear congruential generator*. The case of
$c=0$ is called a *multiplicative* congruential generator; it is
particularly useful for really fast computations.

For illustrative values of $a$ and $m$, Microsoft’s Visual Basic uses
$m=2^{24}$, $a=1,140,671,485$, and $c = 12,820,163$ (see
<http://support.microsoft.com/kb/231847>). This is the engine underlying
the random number generation in Microsoft’s Excel program.

The sequence used by the analyst is defined as $U_n=B_n/m.$ The analyst
may interpret the sequence {$U_{i}$} to be (approximately) identically
and independently uniformly distributed on the interval (0,1). To
illustrate the algorithm, consider the following.

**Example.** Take $m=15$, $a=3$, $c=2$ and $B_0=1$. Then we have:



   step $n$  $B_n$                                      $U_n$
  ---------- --------------------------------- ------------------------
      0      $B_0=1$                           
      1      $B_1 =\mod(3 \times 1 +2) = 5$      $U_1 = \frac{5}{15}$
      2      $B_2 =\mod(3 \times 5 +2) = 2$      $U_2 = \frac{2}{15}$
      3      $B_3 =\mod(3 \times 2 +2) = 8$     $U_3 = \frac{8}{15}$
      4      $B_4 =\mod(3 \times 8 +2) = 11$    $U_4 = \frac{11}{15}$
  ---------- --------------------------------- ------------------------

Sometimes computer generated random results are known as *pseudo*-random
numbers to reflect the fact that they are machine generated and can be
replicated. That is, despite the fact that {$U_{i}$} appears to be
i.i.d, it can be reproduced by using the same seed number (and the same
algorithm). The ability to replicate results can be a tremendous tool as
you use simulation while trying to uncover patterns in a business
process.

The linear congruential generator is just one method of producing
pseudo-random outcomes. It is easy to understand and is (still) widely
used. The linear congruential generator does have limitations, including
the fact that it is possible to detect long-run patterns over time in
the sequences generated (recall that we can interpret “independence” to
mean a total lack of functional patterns). Not surprisingly, advanced
techniques have been developed that address some of this method’s
drawbacks.

## Inverse Transform

With the sequence of uniform random numbers, we next transform them to a
distribution of interest. Let $F$ represent a distribution function of
interest. Then, use the *inverse transform*
$$X_i=F^{-1}\left( U_i \right) .$$ The result is that the sequence
{$X_{i}$} is approximately i.i.d. with distribution function $F$.

To interpret the result, recall that a distribution function, $F$, is
monotonically increasing and so the inverse function, $F^{-1}$, is
well-defined. The inverse distribution function (also known as the
*quantile function*), is defined as $$\begin{aligned}
F^{-1}(y) = \inf_x \{ F(x) \ge y \} ,\end{aligned}$$ where “$\inf$”
stands for “infimum”, or the greatest lower bound.

**Inverse Transform Visualization.** Here is a graph to help you
visualize the inverse transform. When the random variable is continuous,
the distribution function is strictly increasing and we can readily
identify a unique inverse at each point of the distribution.


```{r InverseDF, fig.cap='Inverse of a Distribution Function', out.width='60%', fig.asp=.75, fig.align='center', echo=FALSE}
plot.new()
par(cex=1.3)
x <- seq(0, 10, by=.01)
y<- 1-exp(-x/6)
plot(x,y, xlim=c(0, 10), ylim=c(0, 1), 
        xlab="", type="l", ylab="",xaxs="i", yaxs="i", xaxt="n", yaxt="n")
y.6 <- 1 -exp(-1)
mtext("y=F(x)", side=2, line=2, cex=1.3, las=2, padj=-4, adj=.5)
axis(1, at=5.2, labels="x=")
axis(1, at=6, labels=expression({F^{-1}} (y)))
segments(6,0,6,y.6)
segments(0,y.6,6,y.6)
axis(1, at=0)
axis(2, at=0)
```

The inverse transform result is available when the underlying random
variable is continuous, discrete or a mixture. Here is a series of
examples to illustrate its scope of applications.

**Exponential Distribution Example.** Suppose that we would like to
generate observations from an exponential distribution with scale
parameter $\theta$ so that $F(x) = 1 - e^{-x/\theta}$. To compute the
inverse transform, we can use the following steps: $$\begin{aligned}
 y = F(x) &\Leftrightarrow  y = 1-e^{-x/\theta} \\
  &\Leftrightarrow -\theta \ln(1-y) = x = F^{-1}(y) .\end{aligned}$$
Thus, if $U$ has a uniform (0,1) distribution, then
$X = -\theta \ln(1-U)$ has an exponential distribution with parameter
$\theta$.

*Some Numbers.* Take $\theta = 10$ and generate three random numbers to
get

  ------------------- ------------ ------------- -------------
  $U$                   0.26321364   0.196884752   0.897884218
  $X = -10\ln(1-U)$     1.32658423   0.952221285   9.909071325
  ------------------- ------------ ------------- -------------

**Pareto Distribution Example.** Suppose that we would like to generate
observations from a Pareto distribution with parameters $\alpha$ and
$\theta$ so that
$F(x) = 1 - \left(\frac{\theta}{x+\theta} \right)^{\alpha}$. To compute
the inverse transform, we can use the following steps: 
$$\begin{aligned}
 y = F(x) &\Leftrightarrow 1-y = \left(\frac{\theta}{x+\theta} \right)^{\alpha} \\
  &\Leftrightarrow \left(1-y\right)^{-1/\alpha} = \frac{x+\theta}{\theta} = \frac{x}{\theta} +1 \\
    &\Leftrightarrow \theta \left((1-y)^{-1/\alpha} - 1\right) = x = F^{-1}(y) .\end{aligned}$$
Thus, $X = \theta \left((1-U)^{-1/\alpha} - 1\right)$ has a Pareto
distribution with parameters $\alpha$ and $\theta$.


**Inverse Transform Justification.** Why does the random variable
$X = F^{-1}(U)$ have a distribution function “$F$”?

This is easy to establish in the continuous case. Because $U$ is a
Uniform random variable on (0,1), we know that $\Pr(U \le y) = y$, for
$0 \le y \le 1$. Thus, $$\begin{aligned}
\Pr(X \le x) &= \Pr(F^{-1}(U) \le x) \\
 &= \Pr(F(F^{-1}(U)) \le F(x)) \\
&= \Pr(U \le F(x)) = F(x)\end{aligned}$$ as required. The key step is
that $ F(F^{-1}(u)) = u$ for each $u$, which is clearly true when $F$ is
strictly increasing.


**Bernoulli Distribution Example.** Suppose that we wish to simulate
random variables from a Bernoulli distribution with parameter $p=0.85$.
A graph of the cumulative distribution function shows that the quantile
function can be written as

```{r BinaryDF, fig.cap='Distribution Function of a Binary Random Variable', out.width='50%', fig.asp=.75, fig.align='center', echo=FALSE}
time = seq(-1,2,0.01)
Ftime = c(rep(0,100),rep(.85,100),rep(1,101))

plot(time,Ftime, ylim=c(0,1), xlab="x",ylab="",pch=19, cex=.2)#,type="l")
mtext("F(x)", side=2, at=1.1, las=1, cex=1.2, adj=1.6)
segments(0,0,0,0.85)#,code=4)
segments(1,0.85,1,1)#,code=4)
symbols(0,.85,circles=.03, add=TRUE,bg="black",inches=FALSE)
symbols(0,0,circles=.03, add=TRUE,inches=FALSE)
symbols(1,.85,circles=.03, add=TRUE,inches=FALSE)
symbols(1,1,circles=.03, add=TRUE,bg="black",inches=FALSE)
```

<!-- ![image](BinaryDF.eps){width=".6\textwidth"} -->

$$\begin{aligned}
F^{-1}(y) = \left\{ \begin{array}{cc}
              0 & 0<y \leq 0.85 \\
              1 & 0.85 < y  \leq  1.0 .
            \end{array} \right.\end{aligned}$$

Thus, with the inverse transform we may define $$\begin{aligned}
X = \left\{ \begin{array}{cc}
              0 & 0<U \leq 0.85  \\
              1 &  0.85 < U  \leq  1.0
            \end{array} \right.\end{aligned}$$ *Some Numbers.* Generate
three random numbers to get

  ----------------- ------------ ------------- -------------
  $U$                 0.26321364   0.196884752   0.897884218
  $X =F^{-1}(U)$              0             0             1
  ----------------- ------------ ------------- -------------



**Discrete Distribution Example.** Consider the time of a machine
failure in the first five years. The distribution of failure times is
given as:

  ------------- ----- ----- ----- ----- -----
  Time ($x$)        1     2     3     4     5
  probability     0.1   0.2   0.1   0.4   0.2
  $F(x)$          0.1   0.3   0.4   0.8   1.0
  ------------- ----- ----- ----- ----- -----

```{r DiscreteDF, fig.cap='Distribution Function of a Discrete Random Variable', out.width='80%', fig.asp=.75, fig.align='center', echo=FALSE}
time = seq(0,6,0.01)
Ftime = c(rep(0,100),rep(0.1,100),rep(0.3,100),rep(0.4,100),
          rep(0.8,100),rep(1,101))
plot(time,Ftime, ylim=c(0,1), xlab="x",ylab="",pch=19, cex=.2)#,type="l")
mtext("F(x)", side=2, at=1.1, las=1, cex=1.2, adj=1.6)
segments(1,0,1,0.1)#,code=4)
segments(2,0.1,2,.3)#,code=4)
segments(3,0.3,3,.4)#,code=4)
segments(4,0.4,4,.8)#,code=4)
segments(5,0.8,5,1)#,code=4)

symbols(1,0,circles=.05, add=TRUE,inches=FALSE)
symbols(1,.1,circles=.05, add=TRUE,bg="black",inches=FALSE)
symbols(2,.1,circles=.05, add=TRUE,inches=FALSE)
symbols(2,.3,circles=.05, add=TRUE,bg="black",inches=FALSE)
symbols(3,.3,circles=.05, add=TRUE,inches=FALSE)
symbols(3,.4,circles=.05, add=TRUE,bg="black",inches=FALSE)
symbols(4,.4,circles=.05, add=TRUE,inches=FALSE)
symbols(4,.8,circles=.05, add=TRUE,bg="black",inches=FALSE)
symbols(5,.8,circles=.05, add=TRUE,inches=FALSE)
symbols(5,1,circles=.05, add=TRUE,bg="black",inches=FALSE)
```

Using the graph of the distribution function, with the inverse transform
we may define
$$\begin{aligned}
X = \left\{ \begin{array}{cc}
              1 &   0<U  \leq 0.1  \\
              2 &  0.1 < U  \leq  0.3\\
              3 &  0.3 < U  \leq  0.4\\
              4 &  0.4 < U  \leq  0.8  \\
              5 &  0.8 < U  \leq  1.0     .
            \end{array} \right.\end{aligned}$$

For general discrete random variables there may not be an ordering of
outcomes. For example, a person could own one of five types of life
insurance products and we might use the following algorithm to generate
random outcomes: 

$$\begin{aligned}
X = \left\{ \begin{array}{cc}
  \textrm{whole life} &   0<U  \leq 0.1  \\
 \textrm{endowment} &  0.1 < U  \leq  0.3\\
\textrm{term life} &  0.3 < U  \leq  0.4\\
  \textrm{universal life} &  0.4 < U  \leq  0.8  \\
  \textrm{variable life} &  0.8 < U  \leq  1.0 .
            \end{array} \right.\end{aligned}$$ 
            
Another analyst may use an alternative procedure such as: 

$$\begin{aligned}
X = \left\{ \begin{array}{cc}
  \textrm{whole life} &   0.9<U<1.0  \\
 \textrm{endowment} &  0.7 \leq U < 0.9\\
\textrm{term life} &  0.6 \leq U < 0.7\\
  \textrm{universal life} &  0.2 \leq U < 0.6  \\
  \textrm{variable life} &  0 \leq U < 0.2 .
            \end{array} \right.\end{aligned}$$
            
            
Both algorithms produce (in the long-run) the same probabilities, e.g., $\Pr(\textrm{whole life})=0.1$, and so forth. So, neither is incorrect. You should be aware that there is “more than one way to skin a cat.” (What an old expression!) Similarly, you could use an alterative
algorithm for ordered outcomes (such as failure times 1, 2, 3, 4, or 5, above).

**Mixed Distribution Example.** Consider a random variable that is 0
with probability 70% and is exponentially distributed with parameter
$\theta= 10,000$ with probability 30%. In practice, this might
correspond to a 70% chance of having no insurance claims and a 30%
chance of a claim - if a claim occurs, then it is exponentially
distributed. The distribution function is given as

$$\begin{aligned}
F(y) = \left\{ \begin{array}{cc}
              0 &  x<0  \\
              1 - 0.3 \exp(-x/10000) & x \ge 0 .
            \end{array} \right.\end{aligned}$$
            

```{r MixedDF, fig.cap='Distribution Function of a Hybrid Random Variable', out.width='60%', fig.asp=.75, fig.align='center', echo=FALSE}
time = seq(-1000,40000,10)
Ftime = 1 - .3*exp(-0.0001*time)
Ftime = Ftime*(time>0)
plot(time,Ftime, ylim=c(0,1), xlab="x",ylab="",pch=19, cex=.2)#,xaxt="n")
#axis(1, at=seq(0,40000, by=10000), font=10, cex=0.005, tck=0.01)
mtext("F(x)", side=2, at=1.1, las=1, cex=1.2, adj=1.6)
segments(0,0,0,0.7)#,code=4)
symbols(0,0,circles=500, add=TRUE,inches=FALSE)
symbols(0,.7,circles=500, add=TRUE,bg="black",inches=FALSE)
```

<!-- ![image](MixedDF.eps){width=".6\textwidth"} -->

From the graph, we can see that the inverse transform for generating
random variables with this distribution function is

$$\begin{aligned}
X = F^{-1}(U) = \left\{ \begin{array}{cc}
              0 &  0< U  \leq  0.7  \\
              -1000 \ln (\frac{1-U}{0.3}) & 0.7 < U < 1 .
            \end{array} \right.\end{aligned}$$

As you have seen, for the discrete and mixed random variables, the key
is to draw a graph of the distribution function that allows you to
visualize potential values of the inverse function.

## How Many Simulated Values?

There are many topics to be described in the study of simulation (and
fortunately many good sources to help you). The best way to appreciate
simulation is to experience it. One topic that inevitably comes up is
the number of simulated trials needed to rid yourself of sampling
variability so that you may focus on patterns of interest.

How many simulated values are recommended? 100? 1,000,000? We can use
the central limit theorem to respond to this question. Suppose that we
wish to use simulation to calculate $\mathrm{E~}h(X)$, where $h(\cdot)$
is some known function. Then, based on $R$ simulations (replications),
we get $ X_1,\ldots,X_R$. From this simulated sample, we calculate a
sample average $$\overline{h}_R=\frac{1}{R}\sum_{i=1}^{R} h(X_i)$$ and a
sample standard deviation
$$s_{h,R}^2 = \frac{1}{R} \sum_{i=1}^{R}\left( h(X_i) -\overline{h}_R
\right) ^2.$$
So, $\overline{h}_R$ is your best estimate of
$\mathrm{E~}h(X)$ and $s_{h,R}^2$ provides an indication of the
uncertainty of your estimate. As one criterion for your confidence in
the result, suppose that you wish to be within 1% of the mean with 95%
certainty. According to the central limit theorem, your estimate should
be approximately normally distributed. Thus, you should continue your
simulation until $$\frac{.01\overline{h}_R}{s_{h,R}/\sqrt{R}}\geq 1.96$$
or equivalently $$R \geq 38,416\frac{s_{h,R}^2}{\overline{h}_R^2}.$$
This criterion is a direct application of the approximate normality
(recall that 1.96 is the 97.5th percentile of the standard normal
curve). Note that $\overline{h}_R$ and $s_{h,R}$ are not known in
advance, so you will have to come up with estimates as you go
(sequentially), either by doing a little pilot study in advance or by
interrupting your procedure intermittently to see if the criterion is
satisfied.


