
```{r include = FALSE}
chapnum = 6

```


#Simulation and Resampling {#C:Simulation}

*Chapter Preview.* Simulation is a computationally intensive method used to solve difficult problems. Instead of creating physical processes and experimenting with them in order to understand their operational characteristics, a simulation study is based on a computer representation - it considers various hypothetical conditions as inputs and summarizes the results. Through simulation, a vast number of hypothetical conditions can be quickly and inexpensively examined. Section \@ref(S:SimulationFundamentals) introduces simulation, a wonderful computational tool that is especially useful in complex, multivariate settings.

We can also use simulation to draw from an empirical distribution - this process is known as resampling. Resampling allows us to assess the uncertainty of estimates in complex models. Section \@ref(S:Bootstrap) introduces resampling in the context of bootstrapping to determine the precision of estimators.

Subsequent sections introduce other topics in resampling. Section \@ref(S:CrossValidation) on cross-validation shows how to use it for model selection and validation. Section \@ref(S:ImportanceSampling) on importance sampling describes resampling in specific regions of interest, such as long-tailed actuarial applications. Section \@ref(S:MCMC) on Monte Carlo Markov Chain (MCMC) introduces the simulation and resampling engine underpinning much of modern Bayesian analysis.


## Simulation Fundamentals {#S:SimulationFundamentals}

***
In this section, you learn how to:

-   Generate approximately independent realizations that are uniformly distributed
-   Transform the uniformly distributed realizations to observations
    from a probability distribution of interest
-   Calculate quantities of interest and determining the precision of the calculated quantities

***

```{r eval =  TRUE, child = './Chapters/SimulationSections/SimulationFundamentals.Rmd'}
```

## Bootstrapping and Resampling {#S:Bootstrap}

***
In this section, you learn how to:

-   Generate a nonparametric bootstrap distribution for a statistic of interest
-   Use the bootstrap distribution to generate estimates of precision for the statistic of interest, including bias, standard deviations, and confidence intervals
-   Perform bootstrap analyses for parametric distributions

***

```{r eval = TRUE, child = 'Chapters/SimulationSections/Bootstrap.Rmd'}
```


## Cross-Validation {#S:CrossValidation}

In this section, you learn how to:

-  Compare and contrast cross-validation to simulation techniques and bootstrap methods.
-  Use cross-validation techniques for model selection
-  Explain the jackknife method as a special case of cross-validation and calculate jackknife estimates of bias and standard errors

***

```{r eval = TRUE, child = './Chapters/SimulationSections/CrossValidation.Rmd'}
```


## Importance Sampling {#S:ImportanceSampling}

```{r eval = TRUE, child = './Chapters/SimulationSections/ImportanceSampling.Rmd'}
```


## Monte Carlo Markov Chain (MCMC)  {#S:MCMC}


***

**This section is being written and is not yet complete nor edited. It is here to give you a flavor of what will be in the final version.**

***


```{r eval = TRUE, child = './Chapters/SimulationSections/MCMC.Rmd'}
```

## Further Resources and Contributors {#Simulation:further-reading-and-resources}

-  Include historical references for jackknife (Quenouille, Tukey, Efron)
-  Here are some links to learn more about [reproducibility and randomness](https://freakonometrics.hypotheses.org/6470) and how to go
[from a random generator to a sample function](https://freakonometrics.hypotheses.org/6638).


####Contributors {-}

- **Arthur Charpentier**, Universit&eacute; du Quebec &aacute; Montreal, and **Edward W. (Jed) Frees**, University of Wisconsin-Madison, are the principal authors of the initial version of this chapter. Email: jfrees@bus.wisc.edu and/or arthur.charpentier@gmail.com for chapter comments and suggested improvements.
- Chapter reviewers include: Write Jed or Arthur; to add you name here.



### TS `r chapnum`.A. Bootstrap Applications in Predictive Modeling

***

**This section is being written.**

***


```{r eval =  FALSE, child = 'SimulationSections/BootstrapPredictModels.Rmd'}
```




